### 2017-01-24 archive/latex/annotated.tex
2017/01/24

\section{Class List}
Here are the classes, structs, unions and interfaces with brief descriptions\+:\begin{DoxyCompactList}
\item\contentsline{section}{\hyperlink{classasync__multi__stream__reader}{async\+\_\+multi\+\_\+stream\+\_\+reader$<$ value\+\_\+type $>$} \\*Main thread read multiple disk files simultaneously }{\pageref{classasync__multi__stream__reader}}{}
\item\contentsline{section}{\hyperlink{classasync__multi__stream__writer}{async\+\_\+multi\+\_\+stream\+\_\+writer$<$ value\+\_\+type $>$} \\*Multi-\/stream writer }{\pageref{classasync__multi__stream__writer}}{}
\item\contentsline{section}{\hyperlink{classasync__stream__reader}{async\+\_\+stream\+\_\+reader$<$ value\+\_\+type $>$} \\*Read data from an external disk file using asynchronous stream }{\pageref{classasync__stream__reader}}{}
\item\contentsline{section}{\hyperlink{classasync__stream__writer}{async\+\_\+stream\+\_\+writer$<$ value\+\_\+type $>$} \\*Write data into a disk file using an asynchronous stream }{\pageref{classasync__stream__writer}}{}
\item\contentsline{section}{\hyperlink{class_basic_i_o}{Basic\+IO} \\*Synchronous read operations }{\pageref{class_basic_i_o}}{}
\item\contentsline{section}{\hyperlink{struct_validate4_1_1_bkt_info}{Validate4$<$ alphabet\+\_\+type, alphabet\+\_\+extension\+\_\+type, size\+\_\+type $>$\+::\+Bkt\+Info} \\*Record L-\/type, S-\/type and L\+MS bucket sizes (no matter empty or non-\/empty) in SA \& L\+CP }{\pageref{struct_validate4_1_1_bkt_info}}{}
\item\contentsline{section}{\hyperlink{structbuffer}{buffer$<$ T $>$} \\*Read buffer }{\pageref{structbuffer}}{}
\item\contentsline{section}{\hyperlink{structbuffer__queue}{buffer\+\_\+queue$<$ buffer\+\_\+type $>$} \\*Queue of buffers }{\pageref{structbuffer__queue}}{}
\item\contentsline{section}{\hyperlink{structchoose__int__types}{choose\+\_\+int\+\_\+types$<$ P\+S $>$} \\*Template for choosing int/uint type according to pointer size }{\pageref{structchoose__int__types}}{}
\item\contentsline{section}{\hyperlink{structchoose__int__types_3_014_01_4}{choose\+\_\+int\+\_\+types$<$ 4 $>$} \\*PS = 4, int32/uint32 }{\pageref{structchoose__int__types_3_014_01_4}}{}
\item\contentsline{section}{\hyperlink{structchoose__int__types_3_018_01_4}{choose\+\_\+int\+\_\+types$<$ 8 $>$} \\*Pointer size = 8, 64-\/bit }{\pageref{structchoose__int__types_3_018_01_4}}{}
\item\contentsline{section}{\hyperlink{struct_ex_tuple_sorter}{Ex\+Tuple\+Sorter$<$ tuple\+\_\+type, tuple\+\_\+comparator\+\_\+type $>$} \\*Template for sorter }{\pageref{struct_ex_tuple_sorter}}{}
\item\contentsline{section}{\hyperlink{struct_ex_vector}{Ex\+Vector$<$ value\+\_\+type $>$} \\*Template for vector }{\pageref{struct_ex_vector}}{}
\item\contentsline{section}{\hyperlink{struct_validate4_1_1_l_m_s_validate}{Validate4$<$ alphabet\+\_\+type, alphabet\+\_\+extension\+\_\+type, size\+\_\+type $>$\+::\+L\+M\+S\+Validate} \\*\hyperlink{class_validate}{Validate} S\+A\+\_\+\+L\+MS \& L\+C\+P\+\_\+\+L\+MS }{\pageref{struct_validate4_1_1_l_m_s_validate}}{}
\item\contentsline{section}{\hyperlink{struct_validate4_1_1_l_scan}{Validate4$<$ alphabet\+\_\+type, alphabet\+\_\+extension\+\_\+type, size\+\_\+type $>$\+::\+L\+Scan} \\*Scan SA \& L\+CP rightward to validate the S\+A-\/values of all the S-\/type suffixes and the L\+C\+P-\/values of them and their left neighbor in SA }{\pageref{struct_validate4_1_1_l_scan}}{}
\item\contentsline{section}{\hyperlink{classstd_1_1numeric__limits_3_01uint128_01_4}{std\+::numeric\+\_\+limits$<$ uint128 $>$} }{\pageref{classstd_1_1numeric__limits_3_01uint128_01_4}}{}
\item\contentsline{section}{\hyperlink{classstd_1_1numeric__limits_3_01uint40_01_4}{std\+::numeric\+\_\+limits$<$ uint40 $>$} }{\pageref{classstd_1_1numeric__limits_3_01uint40_01_4}}{}
\item\contentsline{section}{\hyperlink{structpair}{pair$<$ T1, T2 $>$} \\*Structure pair }{\pageref{structpair}}{}
\item\contentsline{section}{\hyperlink{struct_pair_less1st}{Pair\+Less1st$<$ pair\+\_\+type $>$} \\*Sort pair by 1st component in ascending order }{\pageref{struct_pair_less1st}}{}
\item\contentsline{section}{\hyperlink{struct_pair_less2nd}{Pair\+Less2nd$<$ pair\+\_\+type $>$} \\*Sort pair by 2nd component in ascending order }{\pageref{struct_pair_less2nd}}{}
\item\contentsline{section}{\hyperlink{structquadruple}{quadruple$<$ T1, T2, T3, T4 $>$} \\*Structure quadruple }{\pageref{structquadruple}}{}
\item\contentsline{section}{\hyperlink{struct_validate4_1_1_retrieve_pre}{Validate4$<$ alphabet\+\_\+type, alphabet\+\_\+extension\+\_\+type, size\+\_\+type $>$\+::\+Retrieve\+Pre} \\*Retrieve preceding items (pre\+\_\+ch \& pre\+\_\+t) for L-\/type, S-\/type and L\+MS suffixes }{\pageref{struct_validate4_1_1_retrieve_pre}}{}
\item\contentsline{section}{\hyperlink{struct_validate4_1_1_l_m_s_validate_1_1_r_interval}{Validate4$<$ alphabet\+\_\+type, alphabet\+\_\+extension\+\_\+type, size\+\_\+type $>$\+::\+L\+M\+S\+Validate\+::\+R\+Interval} \\*Store (R \% P)$^\wedge$1, (R \% P)$^\wedge$2, (R \% P)$^\wedge$4, .. }{\pageref{struct_validate4_1_1_l_m_s_validate_1_1_r_interval}}{}
\item\contentsline{section}{\hyperlink{struct_validate_1_1_r_interval}{Validate$<$ alphabet\+\_\+type, size\+\_\+type, offset\+\_\+type $>$\+::\+R\+Interval} \\*Store (R \% P)$^\wedge$1, (R \% P)$^\wedge$2, (R \% P)$^\wedge$4, .. }{\pageref{struct_validate_1_1_r_interval}}{}
\item\contentsline{section}{\hyperlink{struct_validate2_1_1_r_interval}{Validate2$<$ alphabet\+\_\+type, size\+\_\+type $>$\+::\+R\+Interval} \\*Store (R \% P)$^\wedge$1, (R \% P)$^\wedge$2, (R \% P)$^\wedge$4, .. }{\pageref{struct_validate2_1_1_r_interval}}{}
\item\contentsline{section}{\hyperlink{struct_validate3_1_1_r_interval}{Validate3$<$ alphabet\+\_\+type, size\+\_\+type $>$\+::\+R\+Interval} \\*Store (R \% P)$^\wedge$1, (R \% P)$^\wedge$2, (R \% P)$^\wedge$4, .. }{\pageref{struct_validate3_1_1_r_interval}}{}
\item\contentsline{section}{\hyperlink{struct_validate4_1_1_r_m_q}{Validate4$<$ alphabet\+\_\+type, alphabet\+\_\+extension\+\_\+type, size\+\_\+type $>$\+::\+R\+MQ} \\*\hyperlink{struct_validate4_1_1_r_m_q}{R\+MQ} }{\pageref{struct_validate4_1_1_r_m_q}}{}
\item\contentsline{section}{\hyperlink{struct_validate4_1_1_r_scan}{Validate4$<$ alphabet\+\_\+type, alphabet\+\_\+extension\+\_\+type, size\+\_\+type $>$\+::\+R\+Scan} \\*Scan SA rightward to validate the S\+A-\/values of all the L-\/type suffixes and the L\+C\+P-\/values of them and their left neighbors in SA }{\pageref{struct_validate4_1_1_r_scan}}{}
\item\contentsline{section}{\hyperlink{class_test}{Test$<$ alphabet\+\_\+type, alphabet\+\_\+extension\+\_\+type, size\+\_\+type $>$} }{\pageref{class_test}}{}
\item\contentsline{section}{\hyperlink{structtriple}{triple$<$ T1, T2, T3 $>$} \\*Structure triple }{\pageref{structtriple}}{}
\item\contentsline{section}{\hyperlink{structtuple__great__comparator__1st}{tuple\+\_\+great\+\_\+comparator\+\_\+1st$<$ tuple\+\_\+type $>$} \\*Function object for comparing tuples by their first component in ascending order }{\pageref{structtuple__great__comparator__1st}}{}
\item\contentsline{section}{\hyperlink{structtuple__less__comparator__1st}{tuple\+\_\+less\+\_\+comparator\+\_\+1st$<$ tuple\+\_\+type $>$} \\*Function object for comparing tuples by their first component in ascending order }{\pageref{structtuple__less__comparator__1st}}{}
\item\contentsline{section}{\hyperlink{structtuple__less__comparator__2nd}{tuple\+\_\+less\+\_\+comparator\+\_\+2nd$<$ tuple\+\_\+type $>$} \\*Function object for comparing tuples by their first two component in ascending order }{\pageref{structtuple__less__comparator__2nd}}{}
\item\contentsline{section}{\hyperlink{classuint128}{uint128} }{\pageref{classuint128}}{}
\item\contentsline{section}{\hyperlink{classuint40}{uint40} }{\pageref{classuint40}}{}
\item\contentsline{section}{\hyperlink{class_validate}{Validate$<$ alphabet\+\_\+type, size\+\_\+type, offset\+\_\+type $>$} \\*\hyperlink{class_validate}{Validate} suffix and L\+CP arrays using Karp-\/\+Rarbin finger-\/printing function }{\pageref{class_validate}}{}
\item\contentsline{section}{\hyperlink{class_validate2}{Validate2$<$ alphabet\+\_\+type, size\+\_\+type $>$} \\*Suffix and L\+CP array validater }{\pageref{class_validate2}}{}
\item\contentsline{section}{\hyperlink{class_validate3}{Validate3$<$ alphabet\+\_\+type, size\+\_\+type $>$} \\*\hyperlink{class_validate}{Validate} sa and lcp using Karp-\/\+Rabin fingerprinting function }{\pageref{class_validate3}}{}
\item\contentsline{section}{\hyperlink{class_validate4}{Validate4$<$ alphabet\+\_\+type, alphabet\+\_\+extension\+\_\+type, size\+\_\+type $>$} \\*Vaildate SA and L\+CP following the induced-\/sorting principle }{\pageref{class_validate4}}{}
\end{DoxyCompactList}