### 2017-01-24 archive/latex/async__stream__writer_8h.tex

\hypertarget{async__stream__writer_8h}{}\section{code/src/common/async\+\_\+stream\+\_\+writer.h File Reference}
\label{async__stream__writer_8h}\index{code/src/common/async\+\_\+stream\+\_\+writer.\+h@{code/src/common/async\+\_\+stream\+\_\+writer.\+h}}

write elements into disk file using asynchronous I/O operations.

{\ttfamily \#include \char`\"{}common/common.\+h\char`\"{}}\newline
{\ttfamily \#include \char`\"{}basicio.\+h\char`\"{}}\newline
{\ttfamily \#include $<$algorithm$>$}\newline
{\ttfamily \#include $<$condition\+\_\+variable$>$}\newline
{\ttfamily \#include $<$mutex$>$}\newline
{\ttfamily \#include $<$thread$>$}\newline
\subsection*{Classes}
\begin{DoxyCompactItemize}
\item 
struct \hyperlink{structbuffer}{buffer$<$ T $>$}
\begin{DoxyCompactList}\small\item\em read buffer \end{DoxyCompactList}\item 
struct \hyperlink{structbuffer__queue}{buffer\+\_\+queue$<$ buffer\+\_\+type $>$}
\begin{DoxyCompactList}\small\item\em queue of buffers \end{DoxyCompactList}\item 
class \hyperlink{classasync__stream__writer}{async\+\_\+stream\+\_\+writer$<$ value\+\_\+type $>$}
\begin{DoxyCompactList}\small\item\em write data into a disk file using an asynchronous stream \end{DoxyCompactList}\item 
class \hyperlink{classasync__multi__stream__writer}{async\+\_\+multi\+\_\+stream\+\_\+writer$<$ value\+\_\+type $>$}
\begin{DoxyCompactList}\small\item\em multi-\/stream writer \end{DoxyCompactList}\end{DoxyCompactItemize}

\subsection{Detailed Description}
write elements into disk file using asynchronous I/O operations.

Copyright (c) 2016, Sun Yat-\/sen University, All rights reserved

\begin{DoxyAuthor}{Author}
Yi Wu 
\end{DoxyAuthor}
\begin{DoxyDate}{Date}
2016.\+12 
\end{DoxyDate}