### 2017-01-24 archive/latex/basicio_8h.tex

\hypertarget{basicio_8h}{}\section{code/src/common/basicio.h File Reference}
\label{basicio_8h}\index{code/src/common/basicio.\+h@{code/src/common/basicio.\+h}}

provide basic I/O interface for sequential read/write operations

\subsection*{Classes}
\begin{DoxyCompactItemize}
\item 
class \hyperlink{class_basic_i_o}{Basic\+IO}
\begin{DoxyCompactList}\small\item\em synchronous read operations \end{DoxyCompactList}\end{DoxyCompactItemize}

\subsection{Detailed Description}
provide basic I/O interface for sequential read/write operations

Copyright (c) 2016, Sun Yat-\/sen University, All rights reserved

\begin{DoxyAuthor}{Author}
Yi Wu 
\end{DoxyAuthor}
\begin{DoxyDate}{Date}
2016.\+12 
\end{DoxyDate}