### 2017-01-24 archive/latex/classasync__stream__reader.tex

\hypertarget{classasync__stream__reader}{}\section{async\+\_\+stream\+\_\+reader$<$ value\+\_\+type $>$ Class Template Reference}
\label{classasync__stream__reader}\index{async\+\_\+stream\+\_\+reader$<$ value\+\_\+type $>$@{async\+\_\+stream\+\_\+reader$<$ value\+\_\+type $>$}}

read data from an external disk file using asynchronous stream

{\ttfamily \#include $<$async\+\_\+stream\+\_\+reader.\+h$>$}

\subsection*{Public Member Functions}
\begin{DoxyCompactItemize}
\item 
void \hyperlink{classasync__stream__reader_a15c2bd3c259b3a84de866c02dd6c7405}{get\+\_\+full\+\_\+buffer} ()
\begin{DoxyCompactList}\small\item\em get a full buffer \end{DoxyCompactList}\item 
\hyperlink{classasync__stream__reader_a02dc7a7adb24a552a265b30ba8544223}{async\+\_\+stream\+\_\+reader} (const std\+::string \&\+\_\+fn, const \hyperlink{types_8h_a60e8696a4678cd348e991a1f172e53f7}{uint64} \&\+\_\+avail\+\_\+mem=(8ul$<$$<$ 20), const uint64 \&\+\_\+bufnum=4ul)
\begin{DoxyCompactList}\small\item\em ctor, available buffer size is given, skip a number of bytes \end{DoxyCompactList}\item 
value\+\_\+type \hyperlink{classasync__stream__reader_abb382de67d33724ff4fea6714fc67ac8}{read} ()
\begin{DoxyCompactList}\small\item\em get next element in the stream \end{DoxyCompactList}\item 
void \hyperlink{classasync__stream__reader_ac441a02e7bb4708948552ce7ced4e9d4}{read} (value\+\_\+type $\ast$\+\_\+des, \hyperlink{types_8h_a60e8696a4678cd348e991a1f172e53f7}{uint64} \+\_\+num)
\begin{DoxyCompactList}\small\item\em read a sequence of items in the stream \end{DoxyCompactList}\item 
void \hyperlink{classasync__stream__reader_a66bc0e20739b27b9c703be9254cdbe05}{skip} (\hyperlink{types_8h_a60e8696a4678cd348e991a1f172e53f7}{uint64} \+\_\+num)
\begin{DoxyCompactList}\small\item\em skip a sequence of elements in the stream \end{DoxyCompactList}\item 
bool \hyperlink{classasync__stream__reader_accade2559119f3bed538d20f6eec4cfb}{empty} ()
\begin{DoxyCompactList}\small\item\em check if empty \end{DoxyCompactList}\item 
\hyperlink{types_8h_a60e8696a4678cd348e991a1f172e53f7}{uint64} \hyperlink{classasync__stream__reader_a15f6c24ea5065855b1d5df69cc7e3e1a}{bytes\+\_\+read} () const
\begin{DoxyCompactList}\small\item\em get read bytes \end{DoxyCompactList}\item 
\hyperlink{classasync__stream__reader_acfbb53847f349372dd3bb0455d9dfaa9}{$\sim$async\+\_\+stream\+\_\+reader} ()
\begin{DoxyCompactList}\small\item\em destructor \end{DoxyCompactList}\end{DoxyCompactItemize}
\subsection*{Private Types}
\begin{DoxyCompactItemize}
\item 
typedef \hyperlink{structbuffer}{buffer}$<$ value\+\_\+type $>$ \hyperlink{classasync__stream__reader_a48e2b7d24cfceba922a771e9c314b40c}{buffer\+\_\+type}
\item 
typedef \hyperlink{structbuffer__queue}{buffer\+\_\+queue}$<$ value\+\_\+type $>$ \hyperlink{classasync__stream__reader_a5301e887b098a0ade898265ca6935d68}{buffer\+\_\+queue\+\_\+type}
\end{DoxyCompactItemize}
\subsection*{Static Private Member Functions}
\begin{DoxyCompactItemize}
\item 
{\footnotesize template$<$typename T $>$ }\\static void \hyperlink{classasync__stream__reader_aee14e37431d75f5a80831ce6c3632ea2}{io\+\_\+procedure} (\hyperlink{classasync__stream__reader}{async\+\_\+stream\+\_\+reader}$<$ T $>$ $\ast$\+\_\+caller)
\begin{DoxyCompactList}\small\item\em I/O procedure that load data into R\+AM using aysnchronous I/O operations. \end{DoxyCompactList}\end{DoxyCompactItemize}
\subsection*{Private Attributes}
\begin{DoxyCompactItemize}
\item 
std\+::\+F\+I\+LE $\ast$ \hyperlink{classasync__stream__reader_ae071a19d938b35853aecdef137a3cfdb}{m\+\_\+file}
\begin{DoxyCompactList}\small\item\em file pointer \end{DoxyCompactList}\item 
\hyperlink{types_8h_a60e8696a4678cd348e991a1f172e53f7}{uint64} \hyperlink{classasync__stream__reader_a5268366ca42bd2ef430e9e0817e9b274}{m\+\_\+bytes\+\_\+read}
\begin{DoxyCompactList}\small\item\em number of bytes already read \end{DoxyCompactList}\item 
\hyperlink{types_8h_a60e8696a4678cd348e991a1f172e53f7}{uint64} \hyperlink{classasync__stream__reader_a8a1fa1966f7523fd4054e06cdd118234}{m\+\_\+cur\+\_\+buffer\+\_\+pos}
\begin{DoxyCompactList}\small\item\em pos of next element to be read in current buffer \end{DoxyCompactList}\item 
\hyperlink{types_8h_a60e8696a4678cd348e991a1f172e53f7}{uint64} \hyperlink{classasync__stream__reader_a0414700ede611b3cbf26ef640c239709}{m\+\_\+cur\+\_\+buffer\+\_\+filled}
\begin{DoxyCompactList}\small\item\em number of elements in current buffer \end{DoxyCompactList}\item 
\hyperlink{classasync__stream__reader_a48e2b7d24cfceba922a771e9c314b40c}{buffer\+\_\+type} $\ast$ \hyperlink{classasync__stream__reader_a167103fd4cb153047682ec786d234f2f}{m\+\_\+cur\+\_\+buffer}
\begin{DoxyCompactList}\small\item\em pointer to current buffer \end{DoxyCompactList}\item 
std\+::thread $\ast$ \hyperlink{classasync__stream__reader_a50b44a356f5c742f20d7ea31bebc018e}{m\+\_\+io\+\_\+thread}
\begin{DoxyCompactList}\small\item\em io thread \end{DoxyCompactList}\item 
\hyperlink{types_8h_a60e8696a4678cd348e991a1f172e53f7}{uint64} \hyperlink{classasync__stream__reader_acbf67b037e53db6a995079a8e9ffac57}{m\+\_\+items\+\_\+toread}
\begin{DoxyCompactList}\small\item\em number of items to read \end{DoxyCompactList}\item 
\hyperlink{types_8h_a60e8696a4678cd348e991a1f172e53f7}{uint64} \hyperlink{classasync__stream__reader_a81e8aa844cad035565f9c194a0d3968b}{m\+\_\+items\+\_\+read}
\begin{DoxyCompactList}\small\item\em number of items read \end{DoxyCompactList}\item 
\hyperlink{classasync__stream__reader_a5301e887b098a0ade898265ca6935d68}{buffer\+\_\+queue\+\_\+type} $\ast$ \hyperlink{classasync__stream__reader_a15a955920c4b23bd70ef0b9e5aebf0df}{m\+\_\+free\+\_\+buffers}
\begin{DoxyCompactList}\small\item\em queue of free buffers \end{DoxyCompactList}\item 
\hyperlink{classasync__stream__reader_a5301e887b098a0ade898265ca6935d68}{buffer\+\_\+queue\+\_\+type} $\ast$ \hyperlink{classasync__stream__reader_a5713212cfddf2a29459e70a8d2852a44}{m\+\_\+full\+\_\+buffers}
\begin{DoxyCompactList}\small\item\em queue of full buffers \end{DoxyCompactList}\end{DoxyCompactItemize}

\subsection{Detailed Description}
\subsubsection*{template$<$typename value\+\_\+type$>$\newline
class async\+\_\+stream\+\_\+reader$<$ value\+\_\+type $>$}

read data from an external disk file using asynchronous stream

I/O thread continually continually do the following steps\+: (1) retrieve a buffer from empty-\/buffer queue (2) load data into buffer (3) put buffer into full-\/buffer queue main thread read elements as following\+: if current buffer is not finished read, then read the next element; otherwise, put current buffer into empty-\/buffer queue and retrieve a buffer from full-\/buffer queue.

\subsection{Member Typedef Documentation}
\mbox{\Hypertarget{classasync__stream__reader_a5301e887b098a0ade898265ca6935d68}\label{classasync__stream__reader_a5301e887b098a0ade898265ca6935d68}} 
\index{async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}!buffer\+\_\+queue\+\_\+type@{buffer\+\_\+queue\+\_\+type}}
\index{buffer\+\_\+queue\+\_\+type@{buffer\+\_\+queue\+\_\+type}!async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}}
\subsubsection{\texorpdfstring{buffer\+\_\+queue\+\_\+type}{buffer\_queue\_type}}
{\footnotesize\ttfamily template$<$typename value\+\_\+type$>$ \\
typedef \hyperlink{structbuffer__queue}{buffer\+\_\+queue}$<$value\+\_\+type$>$ \hyperlink{classasync__stream__reader}{async\+\_\+stream\+\_\+reader}$<$ value\+\_\+type $>$\+::\hyperlink{classasync__stream__reader_a5301e887b098a0ade898265ca6935d68}{buffer\+\_\+queue\+\_\+type}\hspace{0.3cm}{\ttfamily [private]}}

\mbox{\Hypertarget{classasync__stream__reader_a48e2b7d24cfceba922a771e9c314b40c}\label{classasync__stream__reader_a48e2b7d24cfceba922a771e9c314b40c}} 
\index{async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}!buffer\+\_\+type@{buffer\+\_\+type}}
\index{buffer\+\_\+type@{buffer\+\_\+type}!async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}}
\subsubsection{\texorpdfstring{buffer\+\_\+type}{buffer\_type}}
{\footnotesize\ttfamily template$<$typename value\+\_\+type$>$ \\
typedef \hyperlink{structbuffer}{buffer}$<$value\+\_\+type$>$ \hyperlink{classasync__stream__reader}{async\+\_\+stream\+\_\+reader}$<$ value\+\_\+type $>$\+::\hyperlink{classasync__stream__reader_a48e2b7d24cfceba922a771e9c314b40c}{buffer\+\_\+type}\hspace{0.3cm}{\ttfamily [private]}}

\subsection{Constructor \& Destructor Documentation}
\mbox{\Hypertarget{classasync__stream__reader_a02dc7a7adb24a552a265b30ba8544223}\label{classasync__stream__reader_a02dc7a7adb24a552a265b30ba8544223}} 
\index{async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}!async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}}
\index{async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}!async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}}
\subsubsection{\texorpdfstring{async\+\_\+stream\+\_\+reader()}{async\_stream\_reader()}}
{\footnotesize\ttfamily template$<$typename value\+\_\+type$>$ \\
\hyperlink{classasync__stream__reader}{async\+\_\+stream\+\_\+reader}$<$ value\+\_\+type $>$\+::\hyperlink{classasync__stream__reader}{async\+\_\+stream\+\_\+reader} (\begin{DoxyParamCaption}\item[{const std\+::string \&}]{\+\_\+fn,  }\item[{const \hyperlink{types_8h_a60e8696a4678cd348e991a1f172e53f7}{uint64} \&}]{\+\_\+avail\+\_\+mem = {\ttfamily (8ul~$<$$<$~20)},  }\item[{const \hyperlink{types_8h_a60e8696a4678cd348e991a1f172e53f7}{uint64} \&}]{\+\_\+bufnum = {\ttfamily 4ul} }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

ctor, available buffer size is given, skip a number of bytes

\begin{DoxyParams}{Parameters}
{\em \+\_\+fn} & filename \\
\hline
{\em \+\_\+avail\+\_\+mem} & available memory in bytes \\
\hline
{\em \+\_\+bufnum} & number of buffers \\
\hline
{\em \+\_\+skip\+\_\+bytes} & number of byts to skip \\
\hline
\end{DoxyParams}
\mbox{\Hypertarget{classasync__stream__reader_acfbb53847f349372dd3bb0455d9dfaa9}\label{classasync__stream__reader_acfbb53847f349372dd3bb0455d9dfaa9}} 
\index{async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}!````~async\+\_\+stream\+\_\+reader@{$\sim$async\+\_\+stream\+\_\+reader}}
\index{````~async\+\_\+stream\+\_\+reader@{$\sim$async\+\_\+stream\+\_\+reader}!async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}}
\subsubsection{\texorpdfstring{$\sim$async\+\_\+stream\+\_\+reader()}{~async\_stream\_reader()}}
{\footnotesize\ttfamily template$<$typename value\+\_\+type$>$ \\
\hyperlink{classasync__stream__reader}{async\+\_\+stream\+\_\+reader}$<$ value\+\_\+type $>$\+::$\sim$\hyperlink{classasync__stream__reader}{async\+\_\+stream\+\_\+reader} (\begin{DoxyParamCaption}{ }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

destructor

\subsection{Member Function Documentation}
\mbox{\Hypertarget{classasync__stream__reader_a15f6c24ea5065855b1d5df69cc7e3e1a}\label{classasync__stream__reader_a15f6c24ea5065855b1d5df69cc7e3e1a}} 
\index{async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}!bytes\+\_\+read@{bytes\+\_\+read}}
\index{bytes\+\_\+read@{bytes\+\_\+read}!async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}}
\subsubsection{\texorpdfstring{bytes\+\_\+read()}{bytes\_read()}}
{\footnotesize\ttfamily template$<$typename value\+\_\+type$>$ \\
\hyperlink{types_8h_a60e8696a4678cd348e991a1f172e53f7}{uint64} \hyperlink{classasync__stream__reader}{async\+\_\+stream\+\_\+reader}$<$ value\+\_\+type $>$\+::bytes\+\_\+read (\begin{DoxyParamCaption}{ }\end{DoxyParamCaption}) const\hspace{0.3cm}{\ttfamily [inline]}}

get read bytes

\mbox{\Hypertarget{classasync__stream__reader_accade2559119f3bed538d20f6eec4cfb}\label{classasync__stream__reader_accade2559119f3bed538d20f6eec4cfb}} 
\index{async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}!empty@{empty}}
\index{empty@{empty}!async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}}
\subsubsection{\texorpdfstring{empty()}{empty()}}
{\footnotesize\ttfamily template$<$typename value\+\_\+type$>$ \\
bool \hyperlink{classasync__stream__reader}{async\+\_\+stream\+\_\+reader}$<$ value\+\_\+type $>$\+::empty (\begin{DoxyParamCaption}{ }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

check if empty

\mbox{\Hypertarget{classasync__stream__reader_a15c2bd3c259b3a84de866c02dd6c7405}\label{classasync__stream__reader_a15c2bd3c259b3a84de866c02dd6c7405}} 
\index{async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}!get\+\_\+full\+\_\+buffer@{get\+\_\+full\+\_\+buffer}}
\index{get\+\_\+full\+\_\+buffer@{get\+\_\+full\+\_\+buffer}!async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}}
\subsubsection{\texorpdfstring{get\+\_\+full\+\_\+buffer()}{get\_full\_buffer()}}
{\footnotesize\ttfamily template$<$typename value\+\_\+type$>$ \\
void \hyperlink{classasync__stream__reader}{async\+\_\+stream\+\_\+reader}$<$ value\+\_\+type $>$\+::get\+\_\+full\+\_\+buffer (\begin{DoxyParamCaption}{ }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

get a full buffer

\mbox{\Hypertarget{classasync__stream__reader_aee14e37431d75f5a80831ce6c3632ea2}\label{classasync__stream__reader_aee14e37431d75f5a80831ce6c3632ea2}} 
\index{async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}!io\+\_\+procedure@{io\+\_\+procedure}}
\index{io\+\_\+procedure@{io\+\_\+procedure}!async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}}
\subsubsection{\texorpdfstring{io\+\_\+procedure()}{io\_procedure()}}
{\footnotesize\ttfamily template$<$typename value\+\_\+type$>$ \\
template$<$typename T $>$ \\
static void \hyperlink{classasync__stream__reader}{async\+\_\+stream\+\_\+reader}$<$ value\+\_\+type $>$\+::io\+\_\+procedure (\begin{DoxyParamCaption}\item[{\hyperlink{classasync__stream__reader}{async\+\_\+stream\+\_\+reader}$<$ T $>$ $\ast$}]{\+\_\+caller }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}, {\ttfamily [static]}, {\ttfamily [private]}}

I/O procedure that load data into R\+AM using aysnchronous I/O operations.

\mbox{\Hypertarget{classasync__stream__reader_abb382de67d33724ff4fea6714fc67ac8}\label{classasync__stream__reader_abb382de67d33724ff4fea6714fc67ac8}} 
\index{async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}!read@{read}}
\index{read@{read}!async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}}
\subsubsection{\texorpdfstring{read()}{read()}\hspace{0.1cm}{\footnotesize\ttfamily [1/2]}}
{\footnotesize\ttfamily template$<$typename value\+\_\+type$>$ \\
value\+\_\+type \hyperlink{classasync__stream__reader}{async\+\_\+stream\+\_\+reader}$<$ value\+\_\+type $>$\+::read (\begin{DoxyParamCaption}{ }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

get next element in the stream

\begin{DoxyNote}{Note}
check if empty before calling the function 
\end{DoxyNote}
\mbox{\Hypertarget{classasync__stream__reader_ac441a02e7bb4708948552ce7ced4e9d4}\label{classasync__stream__reader_ac441a02e7bb4708948552ce7ced4e9d4}} 
\index{async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}!read@{read}}
\index{read@{read}!async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}}
\subsubsection{\texorpdfstring{read()}{read()}\hspace{0.1cm}{\footnotesize\ttfamily [2/2]}}
{\footnotesize\ttfamily template$<$typename value\+\_\+type$>$ \\
void \hyperlink{classasync__stream__reader}{async\+\_\+stream\+\_\+reader}$<$ value\+\_\+type $>$\+::read (\begin{DoxyParamCaption}\item[{value\+\_\+type $\ast$}]{\+\_\+des,  }\item[{\hyperlink{types_8h_a60e8696a4678cd348e991a1f172e53f7}{uint64}}]{\+\_\+num }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

read a sequence of items in the stream

\begin{DoxyParams}{Parameters}
{\em \+\_\+des} & target container \\
\hline
{\em \+\_\+num} & number of elements to be read \\
\hline
\end{DoxyParams}
\begin{DoxyNote}{Note}
check if empty before calling the function 
\end{DoxyNote}
\mbox{\Hypertarget{classasync__stream__reader_a66bc0e20739b27b9c703be9254cdbe05}\label{classasync__stream__reader_a66bc0e20739b27b9c703be9254cdbe05}} 
\index{async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}!skip@{skip}}
\index{skip@{skip}!async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}}
\subsubsection{\texorpdfstring{skip()}{skip()}}
{\footnotesize\ttfamily template$<$typename value\+\_\+type$>$ \\
void \hyperlink{classasync__stream__reader}{async\+\_\+stream\+\_\+reader}$<$ value\+\_\+type $>$\+::skip (\begin{DoxyParamCaption}\item[{\hyperlink{types_8h_a60e8696a4678cd348e991a1f172e53f7}{uint64}}]{\+\_\+num }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

skip a sequence of elements in the stream

\begin{DoxyParams}{Parameters}
{\em \+\_\+num} & number of elements to be skipped \\
\hline
\end{DoxyParams}

\subsection{Member Data Documentation}
\mbox{\Hypertarget{classasync__stream__reader_a5268366ca42bd2ef430e9e0817e9b274}\label{classasync__stream__reader_a5268366ca42bd2ef430e9e0817e9b274}} 
\index{async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}!m\+\_\+bytes\+\_\+read@{m\+\_\+bytes\+\_\+read}}
\index{m\+\_\+bytes\+\_\+read@{m\+\_\+bytes\+\_\+read}!async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}}
\subsubsection{\texorpdfstring{m\+\_\+bytes\+\_\+read}{m\_bytes\_read}}
{\footnotesize\ttfamily template$<$typename value\+\_\+type$>$ \\
\hyperlink{types_8h_a60e8696a4678cd348e991a1f172e53f7}{uint64} \hyperlink{classasync__stream__reader}{async\+\_\+stream\+\_\+reader}$<$ value\+\_\+type $>$\+::m\+\_\+bytes\+\_\+read\hspace{0.3cm}{\ttfamily [private]}}

number of bytes already read

\mbox{\Hypertarget{classasync__stream__reader_a167103fd4cb153047682ec786d234f2f}\label{classasync__stream__reader_a167103fd4cb153047682ec786d234f2f}} 
\index{async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}!m\+\_\+cur\+\_\+buffer@{m\+\_\+cur\+\_\+buffer}}
\index{m\+\_\+cur\+\_\+buffer@{m\+\_\+cur\+\_\+buffer}!async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}}
\subsubsection{\texorpdfstring{m\+\_\+cur\+\_\+buffer}{m\_cur\_buffer}}
{\footnotesize\ttfamily template$<$typename value\+\_\+type$>$ \\
\hyperlink{classasync__stream__reader_a48e2b7d24cfceba922a771e9c314b40c}{buffer\+\_\+type}$\ast$ \hyperlink{classasync__stream__reader}{async\+\_\+stream\+\_\+reader}$<$ value\+\_\+type $>$\+::m\+\_\+cur\+\_\+buffer\hspace{0.3cm}{\ttfamily [private]}}

pointer to current buffer

\mbox{\Hypertarget{classasync__stream__reader_a0414700ede611b3cbf26ef640c239709}\label{classasync__stream__reader_a0414700ede611b3cbf26ef640c239709}} 
\index{async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}!m\+\_\+cur\+\_\+buffer\+\_\+filled@{m\+\_\+cur\+\_\+buffer\+\_\+filled}}
\index{m\+\_\+cur\+\_\+buffer\+\_\+filled@{m\+\_\+cur\+\_\+buffer\+\_\+filled}!async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}}
\subsubsection{\texorpdfstring{m\+\_\+cur\+\_\+buffer\+\_\+filled}{m\_cur\_buffer\_filled}}
{\footnotesize\ttfamily template$<$typename value\+\_\+type$>$ \\
\hyperlink{types_8h_a60e8696a4678cd348e991a1f172e53f7}{uint64} \hyperlink{classasync__stream__reader}{async\+\_\+stream\+\_\+reader}$<$ value\+\_\+type $>$\+::m\+\_\+cur\+\_\+buffer\+\_\+filled\hspace{0.3cm}{\ttfamily [private]}}

number of elements in current buffer

\mbox{\Hypertarget{classasync__stream__reader_a8a1fa1966f7523fd4054e06cdd118234}\label{classasync__stream__reader_a8a1fa1966f7523fd4054e06cdd118234}} 
\index{async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}!m\+\_\+cur\+\_\+buffer\+\_\+pos@{m\+\_\+cur\+\_\+buffer\+\_\+pos}}
\index{m\+\_\+cur\+\_\+buffer\+\_\+pos@{m\+\_\+cur\+\_\+buffer\+\_\+pos}!async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}}
\subsubsection{\texorpdfstring{m\+\_\+cur\+\_\+buffer\+\_\+pos}{m\_cur\_buffer\_pos}}
{\footnotesize\ttfamily template$<$typename value\+\_\+type$>$ \\
\hyperlink{types_8h_a60e8696a4678cd348e991a1f172e53f7}{uint64} \hyperlink{classasync__stream__reader}{async\+\_\+stream\+\_\+reader}$<$ value\+\_\+type $>$\+::m\+\_\+cur\+\_\+buffer\+\_\+pos\hspace{0.3cm}{\ttfamily [private]}}

pos of next element to be read in current buffer

\mbox{\Hypertarget{classasync__stream__reader_ae071a19d938b35853aecdef137a3cfdb}\label{classasync__stream__reader_ae071a19d938b35853aecdef137a3cfdb}} 
\index{async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}!m\+\_\+file@{m\+\_\+file}}
\index{m\+\_\+file@{m\+\_\+file}!async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}}
\subsubsection{\texorpdfstring{m\+\_\+file}{m\_file}}
{\footnotesize\ttfamily template$<$typename value\+\_\+type$>$ \\
std\+::\+F\+I\+LE$\ast$ \hyperlink{classasync__stream__reader}{async\+\_\+stream\+\_\+reader}$<$ value\+\_\+type $>$\+::m\+\_\+file\hspace{0.3cm}{\ttfamily [private]}}

file pointer

\mbox{\Hypertarget{classasync__stream__reader_a15a955920c4b23bd70ef0b9e5aebf0df}\label{classasync__stream__reader_a15a955920c4b23bd70ef0b9e5aebf0df}} 
\index{async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}!m\+\_\+free\+\_\+buffers@{m\+\_\+free\+\_\+buffers}}
\index{m\+\_\+free\+\_\+buffers@{m\+\_\+free\+\_\+buffers}!async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}}
\subsubsection{\texorpdfstring{m\+\_\+free\+\_\+buffers}{m\_free\_buffers}}
{\footnotesize\ttfamily template$<$typename value\+\_\+type$>$ \\
\hyperlink{classasync__stream__reader_a5301e887b098a0ade898265ca6935d68}{buffer\+\_\+queue\+\_\+type}$\ast$ \hyperlink{classasync__stream__reader}{async\+\_\+stream\+\_\+reader}$<$ value\+\_\+type $>$\+::m\+\_\+free\+\_\+buffers\hspace{0.3cm}{\ttfamily [private]}}

queue of free buffers

\mbox{\Hypertarget{classasync__stream__reader_a5713212cfddf2a29459e70a8d2852a44}\label{classasync__stream__reader_a5713212cfddf2a29459e70a8d2852a44}} 
\index{async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}!m\+\_\+full\+\_\+buffers@{m\+\_\+full\+\_\+buffers}}
\index{m\+\_\+full\+\_\+buffers@{m\+\_\+full\+\_\+buffers}!async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}}
\subsubsection{\texorpdfstring{m\+\_\+full\+\_\+buffers}{m\_full\_buffers}}
{\footnotesize\ttfamily template$<$typename value\+\_\+type$>$ \\
\hyperlink{classasync__stream__reader_a5301e887b098a0ade898265ca6935d68}{buffer\+\_\+queue\+\_\+type}$\ast$ \hyperlink{classasync__stream__reader}{async\+\_\+stream\+\_\+reader}$<$ value\+\_\+type $>$\+::m\+\_\+full\+\_\+buffers\hspace{0.3cm}{\ttfamily [private]}}

queue of full buffers

\mbox{\Hypertarget{classasync__stream__reader_a50b44a356f5c742f20d7ea31bebc018e}\label{classasync__stream__reader_a50b44a356f5c742f20d7ea31bebc018e}} 
\index{async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}!m\+\_\+io\+\_\+thread@{m\+\_\+io\+\_\+thread}}
\index{m\+\_\+io\+\_\+thread@{m\+\_\+io\+\_\+thread}!async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}}
\subsubsection{\texorpdfstring{m\+\_\+io\+\_\+thread}{m\_io\_thread}}
{\footnotesize\ttfamily template$<$typename value\+\_\+type$>$ \\
std\+::thread$\ast$ \hyperlink{classasync__stream__reader}{async\+\_\+stream\+\_\+reader}$<$ value\+\_\+type $>$\+::m\+\_\+io\+\_\+thread\hspace{0.3cm}{\ttfamily [private]}}

io thread

\mbox{\Hypertarget{classasync__stream__reader_a81e8aa844cad035565f9c194a0d3968b}\label{classasync__stream__reader_a81e8aa844cad035565f9c194a0d3968b}} 
\index{async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}!m\+\_\+items\+\_\+read@{m\+\_\+items\+\_\+read}}
\index{m\+\_\+items\+\_\+read@{m\+\_\+items\+\_\+read}!async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}}
\subsubsection{\texorpdfstring{m\+\_\+items\+\_\+read}{m\_items\_read}}
{\footnotesize\ttfamily template$<$typename value\+\_\+type$>$ \\
\hyperlink{types_8h_a60e8696a4678cd348e991a1f172e53f7}{uint64} \hyperlink{classasync__stream__reader}{async\+\_\+stream\+\_\+reader}$<$ value\+\_\+type $>$\+::m\+\_\+items\+\_\+read\hspace{0.3cm}{\ttfamily [private]}}

number of items read

\mbox{\Hypertarget{classasync__stream__reader_acbf67b037e53db6a995079a8e9ffac57}\label{classasync__stream__reader_acbf67b037e53db6a995079a8e9ffac57}} 
\index{async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}!m\+\_\+items\+\_\+toread@{m\+\_\+items\+\_\+toread}}
\index{m\+\_\+items\+\_\+toread@{m\+\_\+items\+\_\+toread}!async\+\_\+stream\+\_\+reader@{async\+\_\+stream\+\_\+reader}}
\subsubsection{\texorpdfstring{m\+\_\+items\+\_\+toread}{m\_items\_toread}}
{\footnotesize\ttfamily template$<$typename value\+\_\+type$>$ \\
\hyperlink{types_8h_a60e8696a4678cd348e991a1f172e53f7}{uint64} \hyperlink{classasync__stream__reader}{async\+\_\+stream\+\_\+reader}$<$ value\+\_\+type $>$\+::m\+\_\+items\+\_\+toread\hspace{0.3cm}{\ttfamily [private]}}

number of items to read

The documentation for this class was generated from the following file\+:\begin{DoxyCompactItemize}
\item 
code/src/common/\hyperlink{async__stream__reader_8h}{async\+\_\+stream\+\_\+reader.\+h}\end{DoxyCompactItemize}