### 2017-01-24 archive/latex/classstd_1_1numeric__limits_3_01uint128_01_4.tex

\hypertarget{classstd_1_1numeric__limits_3_01uint128_01_4}{}\section{std\+:\+:numeric\+\_\+limits$<$ uint128 $>$ Class Template Reference}
\label{classstd_1_1numeric__limits_3_01uint128_01_4}\index{std\+::numeric\+\_\+limits$<$ uint128 $>$@{std\+::numeric\+\_\+limits$<$ uint128 $>$}}

{\ttfamily \#include $<$types.\+h$>$}

\subsection*{Static Public Member Functions}
\begin{DoxyCompactItemize}
\item 
static \hyperlink{classuint128}{uint128} \hyperlink{classstd_1_1numeric__limits_3_01uint128_01_4_adef4934a87fb473dae963876142d84a9}{min} ()
\item 
static \hyperlink{classuint128}{uint128} \hyperlink{classstd_1_1numeric__limits_3_01uint128_01_4_a7093ce47008ca399ec2cd994099baa2a}{max} ()
\end{DoxyCompactItemize}

\subsection{Member Function Documentation}
\mbox{\Hypertarget{classstd_1_1numeric__limits_3_01uint128_01_4_a7093ce47008ca399ec2cd994099baa2a}\label{classstd_1_1numeric__limits_3_01uint128_01_4_a7093ce47008ca399ec2cd994099baa2a}} 
\index{std\+::numeric\+\_\+limits$<$ uint128 $>$@{std\+::numeric\+\_\+limits$<$ uint128 $>$}!max@{max}}
\index{max@{max}!std\+::numeric\+\_\+limits$<$ uint128 $>$@{std\+::numeric\+\_\+limits$<$ uint128 $>$}}
\subsubsection{\texorpdfstring{max()}{max()}}
{\footnotesize\ttfamily static \hyperlink{classuint128}{uint128} std\+::numeric\+\_\+limits$<$ \hyperlink{classuint128}{uint128} $>$\+::max (\begin{DoxyParamCaption}{ }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}, {\ttfamily [static]}}

\mbox{\Hypertarget{classstd_1_1numeric__limits_3_01uint128_01_4_adef4934a87fb473dae963876142d84a9}\label{classstd_1_1numeric__limits_3_01uint128_01_4_adef4934a87fb473dae963876142d84a9}} 
\index{std\+::numeric\+\_\+limits$<$ uint128 $>$@{std\+::numeric\+\_\+limits$<$ uint128 $>$}!min@{min}}
\index{min@{min}!std\+::numeric\+\_\+limits$<$ uint128 $>$@{std\+::numeric\+\_\+limits$<$ uint128 $>$}}
\subsubsection{\texorpdfstring{min()}{min()}}
{\footnotesize\ttfamily static \hyperlink{classuint128}{uint128} std\+::numeric\+\_\+limits$<$ \hyperlink{classuint128}{uint128} $>$\+::min (\begin{DoxyParamCaption}{ }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}, {\ttfamily [static]}}

The documentation for this class was generated from the following file\+:\begin{DoxyCompactItemize}
\item 
code/src/common/\hyperlink{types_8h}{types.\+h}\end{DoxyCompactItemize}