### 2017-01-24 archive/latex/classuint40.tex

\hypertarget{classuint40}{}\section{uint40 Class Reference}
\label{classuint40}\index{uint40@{uint40}}

{\ttfamily \#include $<$types.\+h$>$}

\subsection*{Public Types}
\begin{DoxyCompactItemize}
\item 
typedef \hyperlink{types_8h_a8f95e75e58492e87412191fabadd8ca8}{uint32} \hyperlink{classuint40_a92fb8d876ca5cf77d4ed86dae0adc2f2}{low\+\_\+type}
\item 
typedef \hyperlink{types_8h_a115946cb5fc5879545e9ccea096a6031}{uint8} \hyperlink{classuint40_a58afcd6aeb1ea1e103b8e85eb6aefec4}{high\+\_\+type}
\end{DoxyCompactItemize}
\subsection*{Public Member Functions}
\begin{DoxyCompactItemize}
\item 
\hyperlink{classuint40_a457b73a04d5f5b2833155af392aaac96}{uint40} ()
\item 
\hyperlink{classuint40_a06f45dcc09d7533278c5fefbf7085440}{uint40} (std\+::uint32\+\_\+t l, std\+::uint8\+\_\+t h)
\item 
\hyperlink{classuint40_ab359a2f13e4d14944ae6ba3245d71ef3}{uint40} (const \hyperlink{classuint40}{uint40} \&a)
\item 
\hyperlink{classuint40_ad4165899ffb9045026ba833e5b8bc347}{uint40} (const std\+::int32\+\_\+t \&a)
\item 
\hyperlink{classuint40_a18e730effdedf9cba7e5524e9aef2778}{uint40} (const std\+::uint32\+\_\+t \&a)
\item 
\hyperlink{classuint40_afde52c521f80c8d54ff9bf600fa12aee}{uint40} (const std\+::uint64\+\_\+t \&a)
\item 
\hyperlink{classuint40_a39ca6ed5905712c2fc9f0f8eefc16989}{uint40} (const std\+::int64\+\_\+t \&a)
\item 
void \hyperlink{classuint40_a4d8e0bfd4cb5f959d89d3966c10a28bc}{set\+\_\+high} (const \hyperlink{types_8h_a115946cb5fc5879545e9ccea096a6031}{uint8} \&\+\_\+high)
\item 
void \hyperlink{classuint40_a451e74374aa8f180e94d341e18352d4c}{set\+\_\+low} (const \hyperlink{types_8h_a8f95e75e58492e87412191fabadd8ca8}{uint32} \&\+\_\+low)
\item 
void \hyperlink{classuint40_ab23ea17f6248ac0e7dbacbb71a96179b}{set} (const \hyperlink{types_8h_a8f95e75e58492e87412191fabadd8ca8}{uint32} \&\+\_\+low, const \hyperlink{types_8h_a115946cb5fc5879545e9ccea096a6031}{uint8} \+\_\+high)
\item 
\hyperlink{types_8h_a8f95e75e58492e87412191fabadd8ca8}{uint32} \hyperlink{classuint40_a0037fdf484bb069add9d9ee8c65dbc17}{get\+\_\+low} ()
\item 
\hyperlink{types_8h_a115946cb5fc5879545e9ccea096a6031}{uint8} \hyperlink{classuint40_ac08ecc53d6ae5d245218a768c3ac105c}{get\+\_\+high} ()
\item 
\hyperlink{classuint40_afca52041b179c0002070c33e7dd101a9}{operator uint64\+\_\+t} () const
\item 
bool \hyperlink{classuint40_afa6fa84dfd61e527efa6e4457abc2d81}{operator==} (const \hyperlink{classuint40}{uint40} \&b) const
\item 
bool \hyperlink{classuint40_af17f2004231903079eb4e36b64c96180}{operator!=} (const \hyperlink{classuint40}{uint40} \&b) const
\end{DoxyCompactItemize}
\subsection*{Private Attributes}
\begin{DoxyCompactItemize}
\item 
\hyperlink{classuint40_a92fb8d876ca5cf77d4ed86dae0adc2f2}{low\+\_\+type} \hyperlink{classuint40_a216fde2e6997b39cb34b32daa869cef3}{low}
\item 
\hyperlink{classuint40_a58afcd6aeb1ea1e103b8e85eb6aefec4}{high\+\_\+type} \hyperlink{classuint40_a830b04aa389a4e72c446cc88de38b656}{high}
\end{DoxyCompactItemize}

\subsection{Member Typedef Documentation}
\mbox{\Hypertarget{classuint40_a58afcd6aeb1ea1e103b8e85eb6aefec4}\label{classuint40_a58afcd6aeb1ea1e103b8e85eb6aefec4}} 
\index{uint40@{uint40}!high\+\_\+type@{high\+\_\+type}}
\index{high\+\_\+type@{high\+\_\+type}!uint40@{uint40}}
\subsubsection{\texorpdfstring{high\+\_\+type}{high\_type}}
{\footnotesize\ttfamily typedef \hyperlink{types_8h_a115946cb5fc5879545e9ccea096a6031}{uint8} \hyperlink{classuint40_a58afcd6aeb1ea1e103b8e85eb6aefec4}{uint40\+::high\+\_\+type}}

\mbox{\Hypertarget{classuint40_a92fb8d876ca5cf77d4ed86dae0adc2f2}\label{classuint40_a92fb8d876ca5cf77d4ed86dae0adc2f2}} 
\index{uint40@{uint40}!low\+\_\+type@{low\+\_\+type}}
\index{low\+\_\+type@{low\+\_\+type}!uint40@{uint40}}
\subsubsection{\texorpdfstring{low\+\_\+type}{low\_type}}
{\footnotesize\ttfamily typedef \hyperlink{types_8h_a8f95e75e58492e87412191fabadd8ca8}{uint32} \hyperlink{classuint40_a92fb8d876ca5cf77d4ed86dae0adc2f2}{uint40\+::low\+\_\+type}}

\subsection{Constructor \& Destructor Documentation}
\mbox{\Hypertarget{classuint40_a457b73a04d5f5b2833155af392aaac96}\label{classuint40_a457b73a04d5f5b2833155af392aaac96}} 
\index{uint40@{uint40}!uint40@{uint40}}
\index{uint40@{uint40}!uint40@{uint40}}
\subsubsection{\texorpdfstring{uint40()}{uint40()}\hspace{0.1cm}{\footnotesize\ttfamily [1/7]}}
{\footnotesize\ttfamily uint40\+::uint40 (\begin{DoxyParamCaption}{ }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

\mbox{\Hypertarget{classuint40_a06f45dcc09d7533278c5fefbf7085440}\label{classuint40_a06f45dcc09d7533278c5fefbf7085440}} 
\index{uint40@{uint40}!uint40@{uint40}}
\index{uint40@{uint40}!uint40@{uint40}}
\subsubsection{\texorpdfstring{uint40()}{uint40()}\hspace{0.1cm}{\footnotesize\ttfamily [2/7]}}
{\footnotesize\ttfamily uint40\+::uint40 (\begin{DoxyParamCaption}\item[{std\+::uint32\+\_\+t}]{l,  }\item[{std\+::uint8\+\_\+t}]{h }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

\mbox{\Hypertarget{classuint40_ab359a2f13e4d14944ae6ba3245d71ef3}\label{classuint40_ab359a2f13e4d14944ae6ba3245d71ef3}} 
\index{uint40@{uint40}!uint40@{uint40}}
\index{uint40@{uint40}!uint40@{uint40}}
\subsubsection{\texorpdfstring{uint40()}{uint40()}\hspace{0.1cm}{\footnotesize\ttfamily [3/7]}}
{\footnotesize\ttfamily uint40\+::uint40 (\begin{DoxyParamCaption}\item[{const \hyperlink{classuint40}{uint40} \&}]{a }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

\mbox{\Hypertarget{classuint40_ad4165899ffb9045026ba833e5b8bc347}\label{classuint40_ad4165899ffb9045026ba833e5b8bc347}} 
\index{uint40@{uint40}!uint40@{uint40}}
\index{uint40@{uint40}!uint40@{uint40}}
\subsubsection{\texorpdfstring{uint40()}{uint40()}\hspace{0.1cm}{\footnotesize\ttfamily [4/7]}}
{\footnotesize\ttfamily uint40\+::uint40 (\begin{DoxyParamCaption}\item[{const std\+::int32\+\_\+t \&}]{a }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

\mbox{\Hypertarget{classuint40_a18e730effdedf9cba7e5524e9aef2778}\label{classuint40_a18e730effdedf9cba7e5524e9aef2778}} 
\index{uint40@{uint40}!uint40@{uint40}}
\index{uint40@{uint40}!uint40@{uint40}}
\subsubsection{\texorpdfstring{uint40()}{uint40()}\hspace{0.1cm}{\footnotesize\ttfamily [5/7]}}
{\footnotesize\ttfamily uint40\+::uint40 (\begin{DoxyParamCaption}\item[{const std\+::uint32\+\_\+t \&}]{a }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

\mbox{\Hypertarget{classuint40_afde52c521f80c8d54ff9bf600fa12aee}\label{classuint40_afde52c521f80c8d54ff9bf600fa12aee}} 
\index{uint40@{uint40}!uint40@{uint40}}
\index{uint40@{uint40}!uint40@{uint40}}
\subsubsection{\texorpdfstring{uint40()}{uint40()}\hspace{0.1cm}{\footnotesize\ttfamily [6/7]}}
{\footnotesize\ttfamily uint40\+::uint40 (\begin{DoxyParamCaption}\item[{const std\+::uint64\+\_\+t \&}]{a }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

\mbox{\Hypertarget{classuint40_a39ca6ed5905712c2fc9f0f8eefc16989}\label{classuint40_a39ca6ed5905712c2fc9f0f8eefc16989}} 
\index{uint40@{uint40}!uint40@{uint40}}
\index{uint40@{uint40}!uint40@{uint40}}
\subsubsection{\texorpdfstring{uint40()}{uint40()}\hspace{0.1cm}{\footnotesize\ttfamily [7/7]}}
{\footnotesize\ttfamily uint40\+::uint40 (\begin{DoxyParamCaption}\item[{const std\+::int64\+\_\+t \&}]{a }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

\subsection{Member Function Documentation}
\mbox{\Hypertarget{classuint40_ac08ecc53d6ae5d245218a768c3ac105c}\label{classuint40_ac08ecc53d6ae5d245218a768c3ac105c}} 
\index{uint40@{uint40}!get\+\_\+high@{get\+\_\+high}}
\index{get\+\_\+high@{get\+\_\+high}!uint40@{uint40}}
\subsubsection{\texorpdfstring{get\+\_\+high()}{get\_high()}}
{\footnotesize\ttfamily \hyperlink{types_8h_a115946cb5fc5879545e9ccea096a6031}{uint8} uint40\+::get\+\_\+high (\begin{DoxyParamCaption}{ }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

\mbox{\Hypertarget{classuint40_a0037fdf484bb069add9d9ee8c65dbc17}\label{classuint40_a0037fdf484bb069add9d9ee8c65dbc17}} 
\index{uint40@{uint40}!get\+\_\+low@{get\+\_\+low}}
\index{get\+\_\+low@{get\+\_\+low}!uint40@{uint40}}
\subsubsection{\texorpdfstring{get\+\_\+low()}{get\_low()}}
{\footnotesize\ttfamily \hyperlink{types_8h_a8f95e75e58492e87412191fabadd8ca8}{uint32} uint40\+::get\+\_\+low (\begin{DoxyParamCaption}{ }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

\mbox{\Hypertarget{classuint40_afca52041b179c0002070c33e7dd101a9}\label{classuint40_afca52041b179c0002070c33e7dd101a9}} 
\index{uint40@{uint40}!operator uint64\+\_\+t@{operator uint64\+\_\+t}}
\index{operator uint64\+\_\+t@{operator uint64\+\_\+t}!uint40@{uint40}}
\subsubsection{\texorpdfstring{operator uint64\+\_\+t()}{operator uint64\_t()}}
{\footnotesize\ttfamily uint40\+::operator uint64\+\_\+t (\begin{DoxyParamCaption}{ }\end{DoxyParamCaption}) const\hspace{0.3cm}{\ttfamily [inline]}}

\mbox{\Hypertarget{classuint40_af17f2004231903079eb4e36b64c96180}\label{classuint40_af17f2004231903079eb4e36b64c96180}} 
\index{uint40@{uint40}!operator"!=@{operator"!=}}
\index{operator"!=@{operator"!=}!uint40@{uint40}}
\subsubsection{\texorpdfstring{operator"!=()}{operator!=()}}
{\footnotesize\ttfamily bool uint40\+::operator!= (\begin{DoxyParamCaption}\item[{const \hyperlink{classuint40}{uint40} \&}]{b }\end{DoxyParamCaption}) const\hspace{0.3cm}{\ttfamily [inline]}}

\mbox{\Hypertarget{classuint40_afa6fa84dfd61e527efa6e4457abc2d81}\label{classuint40_afa6fa84dfd61e527efa6e4457abc2d81}} 
\index{uint40@{uint40}!operator==@{operator==}}
\index{operator==@{operator==}!uint40@{uint40}}
\subsubsection{\texorpdfstring{operator==()}{operator==()}}
{\footnotesize\ttfamily bool uint40\+::operator== (\begin{DoxyParamCaption}\item[{const \hyperlink{classuint40}{uint40} \&}]{b }\end{DoxyParamCaption}) const\hspace{0.3cm}{\ttfamily [inline]}}

\mbox{\Hypertarget{classuint40_ab23ea17f6248ac0e7dbacbb71a96179b}\label{classuint40_ab23ea17f6248ac0e7dbacbb71a96179b}} 
\index{uint40@{uint40}!set@{set}}
\index{set@{set}!uint40@{uint40}}
\subsubsection{\texorpdfstring{set()}{set()}}
{\footnotesize\ttfamily void uint40\+::set (\begin{DoxyParamCaption}\item[{const \hyperlink{types_8h_a8f95e75e58492e87412191fabadd8ca8}{uint32} \&}]{\+\_\+low,  }\item[{const \hyperlink{types_8h_a115946cb5fc5879545e9ccea096a6031}{uint8}}]{\+\_\+high }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

\mbox{\Hypertarget{classuint40_a4d8e0bfd4cb5f959d89d3966c10a28bc}\label{classuint40_a4d8e0bfd4cb5f959d89d3966c10a28bc}} 
\index{uint40@{uint40}!set\+\_\+high@{set\+\_\+high}}
\index{set\+\_\+high@{set\+\_\+high}!uint40@{uint40}}
\subsubsection{\texorpdfstring{set\+\_\+high()}{set\_high()}}
{\footnotesize\ttfamily void uint40\+::set\+\_\+high (\begin{DoxyParamCaption}\item[{const \hyperlink{types_8h_a115946cb5fc5879545e9ccea096a6031}{uint8} \&}]{\+\_\+high }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

\mbox{\Hypertarget{classuint40_a451e74374aa8f180e94d341e18352d4c}\label{classuint40_a451e74374aa8f180e94d341e18352d4c}} 
\index{uint40@{uint40}!set\+\_\+low@{set\+\_\+low}}
\index{set\+\_\+low@{set\+\_\+low}!uint40@{uint40}}
\subsubsection{\texorpdfstring{set\+\_\+low()}{set\_low()}}
{\footnotesize\ttfamily void uint40\+::set\+\_\+low (\begin{DoxyParamCaption}\item[{const \hyperlink{types_8h_a8f95e75e58492e87412191fabadd8ca8}{uint32} \&}]{\+\_\+low }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

\subsection{Member Data Documentation}
\mbox{\Hypertarget{classuint40_a830b04aa389a4e72c446cc88de38b656}\label{classuint40_a830b04aa389a4e72c446cc88de38b656}} 
\index{uint40@{uint40}!high@{high}}
\index{high@{high}!uint40@{uint40}}
\subsubsection{\texorpdfstring{high}{high}}
{\footnotesize\ttfamily \hyperlink{classuint40_a58afcd6aeb1ea1e103b8e85eb6aefec4}{high\+\_\+type} uint40\+::high\hspace{0.3cm}{\ttfamily [private]}}

\mbox{\Hypertarget{classuint40_a216fde2e6997b39cb34b32daa869cef3}\label{classuint40_a216fde2e6997b39cb34b32daa869cef3}} 
\index{uint40@{uint40}!low@{low}}
\index{low@{low}!uint40@{uint40}}
\subsubsection{\texorpdfstring{low}{low}}
{\footnotesize\ttfamily \hyperlink{classuint40_a92fb8d876ca5cf77d4ed86dae0adc2f2}{low\+\_\+type} uint40\+::low\hspace{0.3cm}{\ttfamily [private]}}

The documentation for this class was generated from the following file\+:\begin{DoxyCompactItemize}
\item 
code/src/common/\hyperlink{types_8h}{types.\+h}\end{DoxyCompactItemize}