### 2017-01-24 archive/latex/dir_68267d1309a1af8e8297ef4c3efbcdba.tex

\hypertarget{dir_68267d1309a1af8e8297ef4c3efbcdba}{}\section{src Directory Reference}
\label{dir_68267d1309a1af8e8297ef4c3efbcdba}\index{src Directory Reference@{src Directory Reference}}
\subsection*{Directories}
\begin{DoxyCompactItemize}
\item 
directory \hyperlink{dir_fdedb0aba14d44ce9d99bc100e026e6a}{common}
\end{DoxyCompactItemize}
\subsection*{Files}
\begin{DoxyCompactItemize}
\item 
file \hyperlink{test_8h}{test.\+h}
\item 
file \hyperlink{validate_8cpp}{validate.\+cpp}
\item 
file \hyperlink{validate_8h}{validate.\+h}
\begin{DoxyCompactList}\small\item\em first method for validating suffix and L\+CP arrays \end{DoxyCompactList}\item 
file \hyperlink{validate2_8cpp}{validate2.\+cpp}
\item 
file \hyperlink{validate2_8h}{validate2.\+h}
\begin{DoxyCompactList}\small\item\em validate sa and lcp using internal memory sorts and Karp-\/rabin finger-\/print function. \end{DoxyCompactList}\item 
file \hyperlink{validate3_8cpp}{validate3.\+cpp}
\item 
file \hyperlink{validate3_8h}{validate3.\+h}
\begin{DoxyCompactList}\small\item\em validate sa and lcp using external memory sorts and Karp-\/rabin finger-\/print function. \end{DoxyCompactList}\item 
file \hyperlink{validate4_8cpp}{validate4.\+cpp}
\item 
file \hyperlink{validate4_8h}{validate4.\+h}
\begin{DoxyCompactList}\small\item\em A two-\/phase validator for suffix and L\+CP arrays verification. \end{DoxyCompactList}\end{DoxyCompactItemize}