### 2017-01-24 archive/latex/dir_86633d98051024a4283a8055eb45f008.tex

\hypertarget{dir_86633d98051024a4283a8055eb45f008}{}\section{code/src/common Directory Reference}
\label{dir_86633d98051024a4283a8055eb45f008}\index{code/src/common Directory Reference@{code/src/common Directory Reference}}
\subsection*{Files}
\begin{DoxyCompactItemize}
\item 
file \hyperlink{async__stream__reader_8h}{async\+\_\+stream\+\_\+reader.\+h}
\begin{DoxyCompactList}\small\item\em read elements from disk file using asychronous I/O operations \end{DoxyCompactList}\item 
file \hyperlink{async__stream__writer_8h}{async\+\_\+stream\+\_\+writer.\+h}
\begin{DoxyCompactList}\small\item\em write elements into disk file using asynchronous I/O operations. \end{DoxyCompactList}\item 
file \hyperlink{basicio_8h}{basicio.\+h}
\begin{DoxyCompactList}\small\item\em provide basic I/O interface for sequential read/write operations \end{DoxyCompactList}\item 
file \hyperlink{common_8h}{common.\+h}
\begin{DoxyCompactList}\small\item\em definition of macro, constant and alias for common use \end{DoxyCompactList}\item 
file \hyperlink{namespace_8h}{namespace.\+h}
\item 
file \hyperlink{tuples_8h}{tuples.\+h}
\item 
file \hyperlink{types_8h}{types.\+h}
\begin{DoxyCompactList}\small\item\em definition of namespace \end{DoxyCompactList}\item 
file \hyperlink{widget_8h}{widget.\+h}
\begin{DoxyCompactList}\small\item\em sort elements in external memory \end{DoxyCompactList}\end{DoxyCompactItemize}