### 2017-01-24 archive/latex/files.tex

\section{File List}
Here is a list of all files with brief descriptions\+:\begin{DoxyCompactList}
\item\contentsline{section}{code/src/\hyperlink{test_8h}{test.\+h} }{\pageref{test_8h}}{}
\item\contentsline{section}{code/src/\hyperlink{validate_8cpp}{validate.\+cpp} }{\pageref{validate_8cpp}}{}
\item\contentsline{section}{code/src/\hyperlink{validate_8h}{validate.\+h} \\*First method for validating suffix and L\+CP arrays }{\pageref{validate_8h}}{}
\item\contentsline{section}{code/src/\hyperlink{validate2_8cpp}{validate2.\+cpp} }{\pageref{validate2_8cpp}}{}
\item\contentsline{section}{code/src/\hyperlink{validate2_8h}{validate2.\+h} \\*\hyperlink{class_validate}{Validate} sa and lcp using internal memory sorts and Karp-\/rabin finger-\/print function }{\pageref{validate2_8h}}{}
\item\contentsline{section}{code/src/\hyperlink{validate3_8cpp}{validate3.\+cpp} }{\pageref{validate3_8cpp}}{}
\item\contentsline{section}{code/src/\hyperlink{validate3_8h}{validate3.\+h} \\*\hyperlink{class_validate}{Validate} sa and lcp using external memory sorts and Karp-\/rabin finger-\/print function }{\pageref{validate3_8h}}{}
\item\contentsline{section}{code/src/\hyperlink{validate4_8cpp}{validate4.\+cpp} }{\pageref{validate4_8cpp}}{}
\item\contentsline{section}{code/src/\hyperlink{validate4_8h}{validate4.\+h} \\*A two-\/phase validator for suffix and L\+CP arrays verification }{\pageref{validate4_8h}}{}
\item\contentsline{section}{code/src/common/\hyperlink{async__stream__reader_8h}{async\+\_\+stream\+\_\+reader.\+h} \\*Read elements from disk file using asychronous I/O operations }{\pageref{async__stream__reader_8h}}{}
\item\contentsline{section}{code/src/common/\hyperlink{async__stream__writer_8h}{async\+\_\+stream\+\_\+writer.\+h} \\*Write elements into disk file using asynchronous I/O operations }{\pageref{async__stream__writer_8h}}{}
\item\contentsline{section}{code/src/common/\hyperlink{basicio_8h}{basicio.\+h} \\*Provide basic I/O interface for sequential read/write operations }{\pageref{basicio_8h}}{}
\item\contentsline{section}{code/src/common/\hyperlink{common_8h}{common.\+h} \\*Definition of macro, constant and alias for common use }{\pageref{common_8h}}{}
\item\contentsline{section}{code/src/common/\hyperlink{namespace_8h}{namespace.\+h} }{\pageref{namespace_8h}}{}
\item\contentsline{section}{code/src/common/\hyperlink{tuples_8h}{tuples.\+h} }{\pageref{tuples_8h}}{}
\item\contentsline{section}{code/src/common/\hyperlink{types_8h}{types.\+h} \\*Definition of namespace }{\pageref{types_8h}}{}
\item\contentsline{section}{code/src/common/\hyperlink{widget_8h}{widget.\+h} \\*Sort elements in external memory }{\pageref{widget_8h}}{}
\item\contentsline{section}{code/test/\hyperlink{mytest1_8cpp}{mytest1.\+cpp} }{\pageref{mytest1_8cpp}}{}
\end{DoxyCompactList}