### 2017-01-24 archive/latex/index.tex

\hypertarget{index_intro_sec}{}\section{Introduction}\label{index_intro_sec}
The implementations of two methods for validating massive SA and L\+CP arrays. The main files for Methods A and B\+: \hyperlink{validate3_8h}{validate3.\+h} \& \hyperlink{validate3_8cpp}{validate3.\+cpp} and \hyperlink{validate4_8h}{validate4.\+h} \& \hyperlink{validate4_8cpp}{validate4.\+cpp}, respectively.\hypertarget{index_copyright_sec}{}\section{Copyright}\label{index_copyright_sec}
Copyright 2017 by Yi Wu, Sun Yat-\/\+Sen university. Free to use, copy, modify or distribute the software.