### 2017-01-24 archive/latex/namespace_8h.tex

\hypertarget{namespace_8h}{}\section{code/src/common/namespace.h File Reference}
\label{namespace_8h}\index{code/src/common/namespace.\+h@{code/src/common/namespace.\+h}}
\subsection*{Macros}
\begin{DoxyCompactItemize}
\item 
\#define \hyperlink{namespace_8h_a9d80567b3813f4e453d1e9ec2ea08c8f}{N\+A\+M\+E\+S\+P\+A\+C\+E\+\_\+\+C\+O\+R\+E\+\_\+\+B\+EG}~namespace core \{
\item 
\#define \hyperlink{namespace_8h_a867f010187a6a1158481891dfb51273f}{N\+A\+M\+E\+S\+P\+A\+C\+E\+\_\+\+C\+O\+R\+E\+\_\+\+E\+ND}~\}
\item 
\#define \hyperlink{namespace_8h_a84670ad82c41e15277ca235f8616b59e}{N\+A\+M\+E\+S\+P\+A\+C\+E\+\_\+\+U\+T\+I\+L\+I\+T\+Y\+\_\+\+B\+EG}~namespace utility \{
\item 
\#define \hyperlink{namespace_8h_afc10b2ddef8af94c352b9e2f24de67ca}{N\+A\+M\+E\+S\+P\+A\+C\+E\+\_\+\+U\+T\+I\+L\+I\+T\+Y\+\_\+\+E\+ND}~\}
\end{DoxyCompactItemize}

\subsection{Macro Definition Documentation}
\mbox{\Hypertarget{namespace_8h_a9d80567b3813f4e453d1e9ec2ea08c8f}\label{namespace_8h_a9d80567b3813f4e453d1e9ec2ea08c8f}} 
\index{namespace.\+h@{namespace.\+h}!N\+A\+M\+E\+S\+P\+A\+C\+E\+\_\+\+C\+O\+R\+E\+\_\+\+B\+EG@{N\+A\+M\+E\+S\+P\+A\+C\+E\+\_\+\+C\+O\+R\+E\+\_\+\+B\+EG}}
\index{N\+A\+M\+E\+S\+P\+A\+C\+E\+\_\+\+C\+O\+R\+E\+\_\+\+B\+EG@{N\+A\+M\+E\+S\+P\+A\+C\+E\+\_\+\+C\+O\+R\+E\+\_\+\+B\+EG}!namespace.\+h@{namespace.\+h}}
\subsubsection{\texorpdfstring{N\+A\+M\+E\+S\+P\+A\+C\+E\+\_\+\+C\+O\+R\+E\+\_\+\+B\+EG}{NAMESPACE\_CORE\_BEG}}
{\footnotesize\ttfamily \#define N\+A\+M\+E\+S\+P\+A\+C\+E\+\_\+\+C\+O\+R\+E\+\_\+\+B\+EG~namespace core \{}

\mbox{\Hypertarget{namespace_8h_a867f010187a6a1158481891dfb51273f}\label{namespace_8h_a867f010187a6a1158481891dfb51273f}} 
\index{namespace.\+h@{namespace.\+h}!N\+A\+M\+E\+S\+P\+A\+C\+E\+\_\+\+C\+O\+R\+E\+\_\+\+E\+ND@{N\+A\+M\+E\+S\+P\+A\+C\+E\+\_\+\+C\+O\+R\+E\+\_\+\+E\+ND}}
\index{N\+A\+M\+E\+S\+P\+A\+C\+E\+\_\+\+C\+O\+R\+E\+\_\+\+E\+ND@{N\+A\+M\+E\+S\+P\+A\+C\+E\+\_\+\+C\+O\+R\+E\+\_\+\+E\+ND}!namespace.\+h@{namespace.\+h}}
\subsubsection{\texorpdfstring{N\+A\+M\+E\+S\+P\+A\+C\+E\+\_\+\+C\+O\+R\+E\+\_\+\+E\+ND}{NAMESPACE\_CORE\_END}}
{\footnotesize\ttfamily \#define N\+A\+M\+E\+S\+P\+A\+C\+E\+\_\+\+C\+O\+R\+E\+\_\+\+E\+ND~\}}

\mbox{\Hypertarget{namespace_8h_a84670ad82c41e15277ca235f8616b59e}\label{namespace_8h_a84670ad82c41e15277ca235f8616b59e}} 
\index{namespace.\+h@{namespace.\+h}!N\+A\+M\+E\+S\+P\+A\+C\+E\+\_\+\+U\+T\+I\+L\+I\+T\+Y\+\_\+\+B\+EG@{N\+A\+M\+E\+S\+P\+A\+C\+E\+\_\+\+U\+T\+I\+L\+I\+T\+Y\+\_\+\+B\+EG}}
\index{N\+A\+M\+E\+S\+P\+A\+C\+E\+\_\+\+U\+T\+I\+L\+I\+T\+Y\+\_\+\+B\+EG@{N\+A\+M\+E\+S\+P\+A\+C\+E\+\_\+\+U\+T\+I\+L\+I\+T\+Y\+\_\+\+B\+EG}!namespace.\+h@{namespace.\+h}}
\subsubsection{\texorpdfstring{N\+A\+M\+E\+S\+P\+A\+C\+E\+\_\+\+U\+T\+I\+L\+I\+T\+Y\+\_\+\+B\+EG}{NAMESPACE\_UTILITY\_BEG}}
{\footnotesize\ttfamily \#define N\+A\+M\+E\+S\+P\+A\+C\+E\+\_\+\+U\+T\+I\+L\+I\+T\+Y\+\_\+\+B\+EG~namespace utility \{}

\mbox{\Hypertarget{namespace_8h_afc10b2ddef8af94c352b9e2f24de67ca}\label{namespace_8h_afc10b2ddef8af94c352b9e2f24de67ca}} 
\index{namespace.\+h@{namespace.\+h}!N\+A\+M\+E\+S\+P\+A\+C\+E\+\_\+\+U\+T\+I\+L\+I\+T\+Y\+\_\+\+E\+ND@{N\+A\+M\+E\+S\+P\+A\+C\+E\+\_\+\+U\+T\+I\+L\+I\+T\+Y\+\_\+\+E\+ND}}
\index{N\+A\+M\+E\+S\+P\+A\+C\+E\+\_\+\+U\+T\+I\+L\+I\+T\+Y\+\_\+\+E\+ND@{N\+A\+M\+E\+S\+P\+A\+C\+E\+\_\+\+U\+T\+I\+L\+I\+T\+Y\+\_\+\+E\+ND}!namespace.\+h@{namespace.\+h}}
\subsubsection{\texorpdfstring{N\+A\+M\+E\+S\+P\+A\+C\+E\+\_\+\+U\+T\+I\+L\+I\+T\+Y\+\_\+\+E\+ND}{NAMESPACE\_UTILITY\_END}}
{\footnotesize\ttfamily \#define N\+A\+M\+E\+S\+P\+A\+C\+E\+\_\+\+U\+T\+I\+L\+I\+T\+Y\+\_\+\+E\+ND~\}}