### 2017-01-24 archive/latex/struct_ex_tuple_sorter.tex

\hypertarget{struct_ex_tuple_sorter}{}\section{Ex\+Tuple\+Sorter$<$ tuple\+\_\+type, tuple\+\_\+comparator\+\_\+type $>$ Struct Template Reference}
\label{struct_ex_tuple_sorter}\index{Ex\+Tuple\+Sorter$<$ tuple\+\_\+type, tuple\+\_\+comparator\+\_\+type $>$@{Ex\+Tuple\+Sorter$<$ tuple\+\_\+type, tuple\+\_\+comparator\+\_\+type $>$}}

template for sorter

{\ttfamily \#include $<$widget.\+h$>$}

\subsection*{Public Types}
\begin{DoxyCompactItemize}
\item 
typedef stxxl\+::sorter$<$ tuple\+\_\+type, tuple\+\_\+comparator\+\_\+type $>$ \hyperlink{struct_ex_tuple_sorter_adf836f2aaa3d62d311b66eccbf5543d5}{sorter}
\item 
typedef tuple\+\_\+comparator\+\_\+type \hyperlink{struct_ex_tuple_sorter_a876d6fabe768a68b9d730053a619298e}{comparator}
\end{DoxyCompactItemize}

\subsection{Detailed Description}
\subsubsection*{template$<$typename tuple\+\_\+type, typename tuple\+\_\+comparator\+\_\+type$>$\newline
struct Ex\+Tuple\+Sorter$<$ tuple\+\_\+type, tuple\+\_\+comparator\+\_\+type $>$}

template for sorter

\subsection{Member Typedef Documentation}
\mbox{\Hypertarget{struct_ex_tuple_sorter_a876d6fabe768a68b9d730053a619298e}\label{struct_ex_tuple_sorter_a876d6fabe768a68b9d730053a619298e}} 
\index{Ex\+Tuple\+Sorter@{Ex\+Tuple\+Sorter}!comparator@{comparator}}
\index{comparator@{comparator}!Ex\+Tuple\+Sorter@{Ex\+Tuple\+Sorter}}
\subsubsection{\texorpdfstring{comparator}{comparator}}
{\footnotesize\ttfamily template$<$typename tuple\+\_\+type, typename tuple\+\_\+comparator\+\_\+type$>$ \\
typedef tuple\+\_\+comparator\+\_\+type \hyperlink{struct_ex_tuple_sorter}{Ex\+Tuple\+Sorter}$<$ tuple\+\_\+type, tuple\+\_\+comparator\+\_\+type $>$\+::\hyperlink{struct_ex_tuple_sorter_a876d6fabe768a68b9d730053a619298e}{comparator}}

\mbox{\Hypertarget{struct_ex_tuple_sorter_adf836f2aaa3d62d311b66eccbf5543d5}\label{struct_ex_tuple_sorter_adf836f2aaa3d62d311b66eccbf5543d5}} 
\index{Ex\+Tuple\+Sorter@{Ex\+Tuple\+Sorter}!sorter@{sorter}}
\index{sorter@{sorter}!Ex\+Tuple\+Sorter@{Ex\+Tuple\+Sorter}}
\subsubsection{\texorpdfstring{sorter}{sorter}}
{\footnotesize\ttfamily template$<$typename tuple\+\_\+type, typename tuple\+\_\+comparator\+\_\+type$>$ \\
typedef stxxl\+::sorter$<$tuple\+\_\+type, tuple\+\_\+comparator\+\_\+type$>$ \hyperlink{struct_ex_tuple_sorter}{Ex\+Tuple\+Sorter}$<$ tuple\+\_\+type, tuple\+\_\+comparator\+\_\+type $>$\+::\hyperlink{struct_ex_tuple_sorter_adf836f2aaa3d62d311b66eccbf5543d5}{sorter}}

The documentation for this struct was generated from the following file\+:\begin{DoxyCompactItemize}
\item 
code/src/common/\hyperlink{widget_8h}{widget.\+h}\end{DoxyCompactItemize}