### 2017-01-24 archive/latex/struct_ex_vector.tex

\hypertarget{struct_ex_vector}{}\section{Ex\+Vector$<$ value\+\_\+type $>$ Struct Template Reference}
\label{struct_ex_vector}\index{Ex\+Vector$<$ value\+\_\+type $>$@{Ex\+Vector$<$ value\+\_\+type $>$}}

template for vector

{\ttfamily \#include $<$widget.\+h$>$}

\subsection*{Public Types}
\begin{DoxyCompactItemize}
\item 
typedef stxxl\+::\+V\+E\+C\+T\+O\+R\+\_\+\+G\+E\+N\+E\+R\+A\+T\+OR$<$ value\+\_\+type, 8/sizeof(value\+\_\+type)+1, 2, \hyperlink{common_8h_a4b1a49a5378e8e0c9aca7501041c8d6d}{K\+\_\+512} $\ast$sizeof(value\+\_\+type)$>$\+::result \hyperlink{struct_ex_vector_aa45a4c6039859d71d03827b07d265287}{vector}
\end{DoxyCompactItemize}

\subsection{Detailed Description}
\subsubsection*{template$<$typename value\+\_\+type$>$\newline
struct Ex\+Vector$<$ value\+\_\+type $>$}

template for vector

\subsection{Member Typedef Documentation}
\mbox{\Hypertarget{struct_ex_vector_aa45a4c6039859d71d03827b07d265287}\label{struct_ex_vector_aa45a4c6039859d71d03827b07d265287}} 
\index{Ex\+Vector@{Ex\+Vector}!vector@{vector}}
\index{vector@{vector}!Ex\+Vector@{Ex\+Vector}}
\subsubsection{\texorpdfstring{vector}{vector}}
{\footnotesize\ttfamily template$<$typename value\+\_\+type$>$ \\
typedef stxxl\+::\+V\+E\+C\+T\+O\+R\+\_\+\+G\+E\+N\+E\+R\+A\+T\+OR$<$value\+\_\+type, 8 / sizeof(value\+\_\+type) + 1, 2, \hyperlink{common_8h_a4b1a49a5378e8e0c9aca7501041c8d6d}{K\+\_\+512} $\ast$ sizeof(value\+\_\+type)$>$\+::result \hyperlink{struct_ex_vector}{Ex\+Vector}$<$ value\+\_\+type $>$\+::\hyperlink{struct_ex_vector_aa45a4c6039859d71d03827b07d265287}{vector}}

The documentation for this struct was generated from the following file\+:\begin{DoxyCompactItemize}
\item 
code/src/common/\hyperlink{widget_8h}{widget.\+h}\end{DoxyCompactItemize}