### 2017-01-24 archive/latex/struct_pair_less1st.tex

\hypertarget{struct_pair_less1st}{}\section{Pair\+Less1st$<$ pair\+\_\+type $>$ Struct Template Reference}
\label{struct_pair_less1st}\index{Pair\+Less1st$<$ pair\+\_\+type $>$@{Pair\+Less1st$<$ pair\+\_\+type $>$}}

sort pair by 1st component in ascending order

{\ttfamily \#include $<$widget.\+h$>$}

\subsection*{Public Member Functions}
\begin{DoxyCompactItemize}
\item 
bool \hyperlink{struct_pair_less1st_a0c8e1c7ad75317672327e20bdafe56e3}{operator()} (const pair\+\_\+type \&\+\_\+a, const pair\+\_\+type \&\+\_\+b) const
\end{DoxyCompactItemize}

\subsection{Detailed Description}
\subsubsection*{template$<$typename pair\+\_\+type$>$\newline
struct Pair\+Less1st$<$ pair\+\_\+type $>$}

sort pair by 1st component in ascending order

\subsection{Member Function Documentation}
\mbox{\Hypertarget{struct_pair_less1st_a0c8e1c7ad75317672327e20bdafe56e3}\label{struct_pair_less1st_a0c8e1c7ad75317672327e20bdafe56e3}} 
\index{Pair\+Less1st@{Pair\+Less1st}!operator()@{operator()}}
\index{operator()@{operator()}!Pair\+Less1st@{Pair\+Less1st}}
\subsubsection{\texorpdfstring{operator()()}{operator()()}}
{\footnotesize\ttfamily template$<$typename pair\+\_\+type $>$ \\
bool \hyperlink{struct_pair_less1st}{Pair\+Less1st}$<$ pair\+\_\+type $>$\+::operator() (\begin{DoxyParamCaption}\item[{const pair\+\_\+type \&}]{\+\_\+a,  }\item[{const pair\+\_\+type \&}]{\+\_\+b }\end{DoxyParamCaption}) const\hspace{0.3cm}{\ttfamily [inline]}}

The documentation for this struct was generated from the following file\+:\begin{DoxyCompactItemize}
\item 
code/src/common/\hyperlink{widget_8h}{widget.\+h}\end{DoxyCompactItemize}