### 2017-01-24 archive/latex/struct_pair_less2nd.tex

\hypertarget{struct_pair_less2nd}{}\section{Pair\+Less2nd$<$ pair\+\_\+type $>$ Struct Template Reference}
\label{struct_pair_less2nd}\index{Pair\+Less2nd$<$ pair\+\_\+type $>$@{Pair\+Less2nd$<$ pair\+\_\+type $>$}}

sort pair by 2nd component in ascending order

{\ttfamily \#include $<$widget.\+h$>$}

\subsection*{Public Member Functions}
\begin{DoxyCompactItemize}
\item 
bool \hyperlink{struct_pair_less2nd_ab794f9af084ecf20527e87634b669f17}{operator()} (const pair\+\_\+type \&\+\_\+a, const pair\+\_\+type \&\+\_\+b) const
\end{DoxyCompactItemize}

\subsection{Detailed Description}
\subsubsection*{template$<$typename pair\+\_\+type$>$\newline
struct Pair\+Less2nd$<$ pair\+\_\+type $>$}

sort pair by 2nd component in ascending order

\subsection{Member Function Documentation}
\mbox{\Hypertarget{struct_pair_less2nd_ab794f9af084ecf20527e87634b669f17}\label{struct_pair_less2nd_ab794f9af084ecf20527e87634b669f17}} 
\index{Pair\+Less2nd@{Pair\+Less2nd}!operator()@{operator()}}
\index{operator()@{operator()}!Pair\+Less2nd@{Pair\+Less2nd}}
\subsubsection{\texorpdfstring{operator()()}{operator()()}}
{\footnotesize\ttfamily template$<$typename pair\+\_\+type $>$ \\
bool \hyperlink{struct_pair_less2nd}{Pair\+Less2nd}$<$ pair\+\_\+type $>$\+::operator() (\begin{DoxyParamCaption}\item[{const pair\+\_\+type \&}]{\+\_\+a,  }\item[{const pair\+\_\+type \&}]{\+\_\+b }\end{DoxyParamCaption}) const\hspace{0.3cm}{\ttfamily [inline]}}

The documentation for this struct was generated from the following file\+:\begin{DoxyCompactItemize}
\item 
code/src/common/\hyperlink{widget_8h}{widget.\+h}\end{DoxyCompactItemize}