### 2017-01-24 archive/latex/struct_validate2_1_1_r_interval.tex

\hypertarget{struct_validate2_1_1_r_interval}{}\section{Validate2$<$ alphabet\+\_\+type, size\+\_\+type $>$\+:\+:R\+Interval Struct Reference}
\label{struct_validate2_1_1_r_interval}\index{Validate2$<$ alphabet\+\_\+type, size\+\_\+type $>$\+::\+R\+Interval@{Validate2$<$ alphabet\+\_\+type, size\+\_\+type $>$\+::\+R\+Interval}}

store (R \% P)$^\wedge$1, (R \% P)$^\wedge$2, (R \% P)$^\wedge$4, ...

\subsection*{Public Member Functions}
\begin{DoxyCompactItemize}
\item 
\hyperlink{struct_validate2_1_1_r_interval_af2e88047192f7e9703b9faa2298cb5fc}{R\+Interval} (\hyperlink{types_8h_a60e8696a4678cd348e991a1f172e53f7}{uint64} \+\_\+len)
\begin{DoxyCompactList}\small\item\em constructor \end{DoxyCompactList}\item 
\hyperlink{common_8h_a7fdaf8b9b3d2f6ae6b10597a8d3f96ee}{fpa\+\_\+type} \hyperlink{struct_validate2_1_1_r_interval_a84006809e4b3ef58bd7335d2f46a92b1}{compute} (\hyperlink{types_8h_a60e8696a4678cd348e991a1f172e53f7}{uint64} \+\_\+interval)
\begin{DoxyCompactList}\small\item\em get fp\+Interval \end{DoxyCompactList}\end{DoxyCompactItemize}
\subsection*{Public Attributes}
\begin{DoxyCompactItemize}
\item 
\hyperlink{common_8h_a7fdaf8b9b3d2f6ae6b10597a8d3f96ee}{fpa\+\_\+type} $\ast$ \hyperlink{struct_validate2_1_1_r_interval_aee05aa76b959338fd2d1c0833f2e35d7}{m\+\_\+data}
\item 
\hyperlink{types_8h_a60e8696a4678cd348e991a1f172e53f7}{uint64} \hyperlink{struct_validate2_1_1_r_interval_ace46ef0a61dad6c184579e23bfae12ce}{m\+\_\+num}
\end{DoxyCompactItemize}

\subsection{Detailed Description}
\subsubsection*{template$<$typename alphabet\+\_\+type, typename size\+\_\+type$>$\newline
struct Validate2$<$ alphabet\+\_\+type, size\+\_\+type $>$\+::\+R\+Interval}

store (R \% P)$^\wedge$1, (R \% P)$^\wedge$2, (R \% P)$^\wedge$4, ...

\subsection{Constructor \& Destructor Documentation}
\mbox{\Hypertarget{struct_validate2_1_1_r_interval_af2e88047192f7e9703b9faa2298cb5fc}\label{struct_validate2_1_1_r_interval_af2e88047192f7e9703b9faa2298cb5fc}} 
\index{Validate2\+::\+R\+Interval@{Validate2\+::\+R\+Interval}!R\+Interval@{R\+Interval}}
\index{R\+Interval@{R\+Interval}!Validate2\+::\+R\+Interval@{Validate2\+::\+R\+Interval}}
\subsubsection{\texorpdfstring{R\+Interval()}{RInterval()}}
{\footnotesize\ttfamily template$<$typename alphabet\+\_\+type , typename size\+\_\+type $>$ \\
\hyperlink{class_validate2}{Validate2}$<$ alphabet\+\_\+type, size\+\_\+type $>$\+::R\+Interval\+::\+R\+Interval (\begin{DoxyParamCaption}\item[{\hyperlink{types_8h_a60e8696a4678cd348e991a1f172e53f7}{uint64}}]{\+\_\+len }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

constructor

\subsection{Member Function Documentation}
\mbox{\Hypertarget{struct_validate2_1_1_r_interval_a84006809e4b3ef58bd7335d2f46a92b1}\label{struct_validate2_1_1_r_interval_a84006809e4b3ef58bd7335d2f46a92b1}} 
\index{Validate2\+::\+R\+Interval@{Validate2\+::\+R\+Interval}!compute@{compute}}
\index{compute@{compute}!Validate2\+::\+R\+Interval@{Validate2\+::\+R\+Interval}}
\subsubsection{\texorpdfstring{compute()}{compute()}}
{\footnotesize\ttfamily template$<$typename alphabet\+\_\+type , typename size\+\_\+type $>$ \\
\hyperlink{common_8h_a7fdaf8b9b3d2f6ae6b10597a8d3f96ee}{fpa\+\_\+type} \hyperlink{class_validate2}{Validate2}$<$ alphabet\+\_\+type, size\+\_\+type $>$\+::R\+Interval\+::compute (\begin{DoxyParamCaption}\item[{\hyperlink{types_8h_a60e8696a4678cd348e991a1f172e53f7}{uint64}}]{\+\_\+interval }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

get fp\+Interval

\subsection{Member Data Documentation}
\mbox{\Hypertarget{struct_validate2_1_1_r_interval_aee05aa76b959338fd2d1c0833f2e35d7}\label{struct_validate2_1_1_r_interval_aee05aa76b959338fd2d1c0833f2e35d7}} 
\index{Validate2\+::\+R\+Interval@{Validate2\+::\+R\+Interval}!m\+\_\+data@{m\+\_\+data}}
\index{m\+\_\+data@{m\+\_\+data}!Validate2\+::\+R\+Interval@{Validate2\+::\+R\+Interval}}
\subsubsection{\texorpdfstring{m\+\_\+data}{m\_data}}
{\footnotesize\ttfamily template$<$typename alphabet\+\_\+type , typename size\+\_\+type $>$ \\
\hyperlink{common_8h_a7fdaf8b9b3d2f6ae6b10597a8d3f96ee}{fpa\+\_\+type}$\ast$ \hyperlink{class_validate2}{Validate2}$<$ alphabet\+\_\+type, size\+\_\+type $>$\+::R\+Interval\+::m\+\_\+data}

\mbox{\Hypertarget{struct_validate2_1_1_r_interval_ace46ef0a61dad6c184579e23bfae12ce}\label{struct_validate2_1_1_r_interval_ace46ef0a61dad6c184579e23bfae12ce}} 
\index{Validate2\+::\+R\+Interval@{Validate2\+::\+R\+Interval}!m\+\_\+num@{m\+\_\+num}}
\index{m\+\_\+num@{m\+\_\+num}!Validate2\+::\+R\+Interval@{Validate2\+::\+R\+Interval}}
\subsubsection{\texorpdfstring{m\+\_\+num}{m\_num}}
{\footnotesize\ttfamily template$<$typename alphabet\+\_\+type , typename size\+\_\+type $>$ \\
\hyperlink{types_8h_a60e8696a4678cd348e991a1f172e53f7}{uint64} \hyperlink{class_validate2}{Validate2}$<$ alphabet\+\_\+type, size\+\_\+type $>$\+::R\+Interval\+::m\+\_\+num}

The documentation for this struct was generated from the following file\+:\begin{DoxyCompactItemize}
\item 
code/src/\hyperlink{validate2_8h}{validate2.\+h}\end{DoxyCompactItemize}