### 2017-01-24 archive/latex/struct_validate4_1_1_r_m_q.tex

\hypertarget{struct_validate4_1_1_r_m_q}{}\section{Validate4$<$ alphabet\+\_\+type, alphabet\+\_\+extension\+\_\+type, size\+\_\+type $>$\+:\+:R\+MQ Struct Reference}
\label{struct_validate4_1_1_r_m_q}\index{Validate4$<$ alphabet\+\_\+type, alphabet\+\_\+extension\+\_\+type, size\+\_\+type $>$\+::\+R\+MQ@{Validate4$<$ alphabet\+\_\+type, alphabet\+\_\+extension\+\_\+type, size\+\_\+type $>$\+::\+R\+MQ}}

\hyperlink{struct_validate4_1_1_r_m_q}{R\+MQ}.

\subsection*{Public Member Functions}
\begin{DoxyCompactItemize}
\item 
\hyperlink{struct_validate4_1_1_r_m_q_a58a58fb442194bba96116dd76b361882}{R\+MQ} ()
\begin{DoxyCompactList}\small\item\em ctor \end{DoxyCompactList}\item 
size\+\_\+type \hyperlink{struct_validate4_1_1_r_m_q_a25c9ec1cab24c2ea2355b8ce828cf108}{get} (const alphabet\+\_\+type \+\_\+ch)
\begin{DoxyCompactList}\small\item\em getter \end{DoxyCompactList}\item 
void \hyperlink{struct_validate4_1_1_r_m_q_a51f01cfbe100d40fc976fd2e1e8b6d16}{update} (size\+\_\+type \+\_\+val)
\begin{DoxyCompactList}\small\item\em update \end{DoxyCompactList}\item 
void \hyperlink{struct_validate4_1_1_r_m_q_a8416e95d49b4b8c05cd814d78017d81f}{reset} (const alphabet\+\_\+type \+\_\+ch)
\begin{DoxyCompactList}\small\item\em reset the range minimum value to val\+\_\+max \end{DoxyCompactList}\end{DoxyCompactItemize}
\subsection*{Private Attributes}
\begin{DoxyCompactItemize}
\item 
const alphabet\+\_\+type \hyperlink{struct_validate4_1_1_r_m_q_a0addfc5d1067fca9a97939554cb0cd70}{ch\+\_\+max}
\item 
const size\+\_\+type \hyperlink{struct_validate4_1_1_r_m_q_a45aa4f738281304a3f6bd5b9d78ef712}{val\+\_\+max}
\item 
std\+::vector$<$ size\+\_\+type $>$ \hyperlink{struct_validate4_1_1_r_m_q_abf525030d892825a6ef20c45eafaae2b}{m\+\_\+rmq}
\end{DoxyCompactItemize}

\subsection{Detailed Description}
\subsubsection*{template$<$typename alphabet\+\_\+type, typename alphabet\+\_\+extension\+\_\+type, typename size\+\_\+type$>$\newline
struct Validate4$<$ alphabet\+\_\+type, alphabet\+\_\+extension\+\_\+type, size\+\_\+type $>$\+::\+R\+MQ}

\hyperlink{struct_validate4_1_1_r_m_q}{R\+MQ}.

Given SA\mbox{[}k\mbox{]} and SA\mbox{[}k + 1\mbox{]} respectively induced from SA\mbox{[}i\mbox{]} and SA\mbox{[}j\mbox{]}, the L\+C\+P-\/value of suf(\+S\+A\mbox{[}k\mbox{]}) and suf(SA\mbox{[}k + 1\mbox{]}) is determined by L\+CP\mbox{[}i + 1, j\mbox{]}. The response to the range-\/minimum query for L\+CP\mbox{[}i + 1, j\mbox{]} returns the minimum value among L\+CP\mbox{[}i + 1\mbox{]}, .., L\+CP\mbox{[}j -\/ 1\mbox{]}, L\+CP\mbox{[}j\mbox{]}.

\subsection{Constructor \& Destructor Documentation}
\mbox{\Hypertarget{struct_validate4_1_1_r_m_q_a58a58fb442194bba96116dd76b361882}\label{struct_validate4_1_1_r_m_q_a58a58fb442194bba96116dd76b361882}} 
\index{Validate4\+::\+R\+MQ@{Validate4\+::\+R\+MQ}!R\+MQ@{R\+MQ}}
\index{R\+MQ@{R\+MQ}!Validate4\+::\+R\+MQ@{Validate4\+::\+R\+MQ}}
\subsubsection{\texorpdfstring{R\+M\+Q()}{RMQ()}}
{\footnotesize\ttfamily template$<$typename alphabet\+\_\+type , typename alphabet\+\_\+extension\+\_\+type , typename size\+\_\+type $>$ \\
\hyperlink{class_validate4}{Validate4}$<$ alphabet\+\_\+type, alphabet\+\_\+extension\+\_\+type, size\+\_\+type $>$\+::R\+M\+Q\+::\+R\+MQ (\begin{DoxyParamCaption}{ }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

ctor

\subsection{Member Function Documentation}
\mbox{\Hypertarget{struct_validate4_1_1_r_m_q_a25c9ec1cab24c2ea2355b8ce828cf108}\label{struct_validate4_1_1_r_m_q_a25c9ec1cab24c2ea2355b8ce828cf108}} 
\index{Validate4\+::\+R\+MQ@{Validate4\+::\+R\+MQ}!get@{get}}
\index{get@{get}!Validate4\+::\+R\+MQ@{Validate4\+::\+R\+MQ}}
\subsubsection{\texorpdfstring{get()}{get()}}
{\footnotesize\ttfamily template$<$typename alphabet\+\_\+type , typename alphabet\+\_\+extension\+\_\+type , typename size\+\_\+type $>$ \\
size\+\_\+type \hyperlink{class_validate4}{Validate4}$<$ alphabet\+\_\+type, alphabet\+\_\+extension\+\_\+type, size\+\_\+type $>$\+::R\+M\+Q\+::get (\begin{DoxyParamCaption}\item[{const alphabet\+\_\+type}]{\+\_\+ch }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

getter

\mbox{\Hypertarget{struct_validate4_1_1_r_m_q_a8416e95d49b4b8c05cd814d78017d81f}\label{struct_validate4_1_1_r_m_q_a8416e95d49b4b8c05cd814d78017d81f}} 
\index{Validate4\+::\+R\+MQ@{Validate4\+::\+R\+MQ}!reset@{reset}}
\index{reset@{reset}!Validate4\+::\+R\+MQ@{Validate4\+::\+R\+MQ}}
\subsubsection{\texorpdfstring{reset()}{reset()}}
{\footnotesize\ttfamily template$<$typename alphabet\+\_\+type , typename alphabet\+\_\+extension\+\_\+type , typename size\+\_\+type $>$ \\
void \hyperlink{class_validate4}{Validate4}$<$ alphabet\+\_\+type, alphabet\+\_\+extension\+\_\+type, size\+\_\+type $>$\+::R\+M\+Q\+::reset (\begin{DoxyParamCaption}\item[{const alphabet\+\_\+type}]{\+\_\+ch }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

reset the range minimum value to val\+\_\+max

\mbox{\Hypertarget{struct_validate4_1_1_r_m_q_a51f01cfbe100d40fc976fd2e1e8b6d16}\label{struct_validate4_1_1_r_m_q_a51f01cfbe100d40fc976fd2e1e8b6d16}} 
\index{Validate4\+::\+R\+MQ@{Validate4\+::\+R\+MQ}!update@{update}}
\index{update@{update}!Validate4\+::\+R\+MQ@{Validate4\+::\+R\+MQ}}
\subsubsection{\texorpdfstring{update()}{update()}}
{\footnotesize\ttfamily template$<$typename alphabet\+\_\+type , typename alphabet\+\_\+extension\+\_\+type , typename size\+\_\+type $>$ \\
void \hyperlink{class_validate4}{Validate4}$<$ alphabet\+\_\+type, alphabet\+\_\+extension\+\_\+type, size\+\_\+type $>$\+::R\+M\+Q\+::update (\begin{DoxyParamCaption}\item[{size\+\_\+type}]{\+\_\+val }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

update

scan all the elements in m\+\_\+rmq, update m\+\_\+rmq\mbox{[}ch\mbox{]} as \+\_\+val if m\+\_\+rmq\mbox{[}ch\mbox{]} $>$ \+\_\+val.

\subsection{Member Data Documentation}
\mbox{\Hypertarget{struct_validate4_1_1_r_m_q_a0addfc5d1067fca9a97939554cb0cd70}\label{struct_validate4_1_1_r_m_q_a0addfc5d1067fca9a97939554cb0cd70}} 
\index{Validate4\+::\+R\+MQ@{Validate4\+::\+R\+MQ}!ch\+\_\+max@{ch\+\_\+max}}
\index{ch\+\_\+max@{ch\+\_\+max}!Validate4\+::\+R\+MQ@{Validate4\+::\+R\+MQ}}
\subsubsection{\texorpdfstring{ch\+\_\+max}{ch\_max}}
{\footnotesize\ttfamily template$<$typename alphabet\+\_\+type , typename alphabet\+\_\+extension\+\_\+type , typename size\+\_\+type $>$ \\
const alphabet\+\_\+type \hyperlink{class_validate4}{Validate4}$<$ alphabet\+\_\+type, alphabet\+\_\+extension\+\_\+type, size\+\_\+type $>$\+::R\+M\+Q\+::ch\+\_\+max\hspace{0.3cm}{\ttfamily [private]}}

\mbox{\Hypertarget{struct_validate4_1_1_r_m_q_abf525030d892825a6ef20c45eafaae2b}\label{struct_validate4_1_1_r_m_q_abf525030d892825a6ef20c45eafaae2b}} 
\index{Validate4\+::\+R\+MQ@{Validate4\+::\+R\+MQ}!m\+\_\+rmq@{m\+\_\+rmq}}
\index{m\+\_\+rmq@{m\+\_\+rmq}!Validate4\+::\+R\+MQ@{Validate4\+::\+R\+MQ}}
\subsubsection{\texorpdfstring{m\+\_\+rmq}{m\_rmq}}
{\footnotesize\ttfamily template$<$typename alphabet\+\_\+type , typename alphabet\+\_\+extension\+\_\+type , typename size\+\_\+type $>$ \\
std\+::vector$<$size\+\_\+type$>$ \hyperlink{class_validate4}{Validate4}$<$ alphabet\+\_\+type, alphabet\+\_\+extension\+\_\+type, size\+\_\+type $>$\+::R\+M\+Q\+::m\+\_\+rmq\hspace{0.3cm}{\ttfamily [private]}}

\mbox{\Hypertarget{struct_validate4_1_1_r_m_q_a45aa4f738281304a3f6bd5b9d78ef712}\label{struct_validate4_1_1_r_m_q_a45aa4f738281304a3f6bd5b9d78ef712}} 
\index{Validate4\+::\+R\+MQ@{Validate4\+::\+R\+MQ}!val\+\_\+max@{val\+\_\+max}}
\index{val\+\_\+max@{val\+\_\+max}!Validate4\+::\+R\+MQ@{Validate4\+::\+R\+MQ}}
\subsubsection{\texorpdfstring{val\+\_\+max}{val\_max}}
{\footnotesize\ttfamily template$<$typename alphabet\+\_\+type , typename alphabet\+\_\+extension\+\_\+type , typename size\+\_\+type $>$ \\
const size\+\_\+type \hyperlink{class_validate4}{Validate4}$<$ alphabet\+\_\+type, alphabet\+\_\+extension\+\_\+type, size\+\_\+type $>$\+::R\+M\+Q\+::val\+\_\+max\hspace{0.3cm}{\ttfamily [private]}}

The documentation for this struct was generated from the following file\+:\begin{DoxyCompactItemize}
\item 
code/src/\hyperlink{validate4_8h}{validate4.\+h}\end{DoxyCompactItemize}