### 2017-01-24 archive/latex/struct_validate4_1_1_retrieve_pre.tex

\hypertarget{struct_validate4_1_1_retrieve_pre}{}\section{Validate4$<$ alphabet\+\_\+type, alphabet\+\_\+extension\+\_\+type, size\+\_\+type $>$\+:\+:Retrieve\+Pre Struct Reference}
\label{struct_validate4_1_1_retrieve_pre}\index{Validate4$<$ alphabet\+\_\+type, alphabet\+\_\+extension\+\_\+type, size\+\_\+type $>$\+::\+Retrieve\+Pre@{Validate4$<$ alphabet\+\_\+type, alphabet\+\_\+extension\+\_\+type, size\+\_\+type $>$\+::\+Retrieve\+Pre}}

retrieve preceding items (pre\+\_\+ch \& pre\+\_\+t) for L-\/type, S-\/type and L\+MS suffixes

\subsection*{Public Member Functions}
\begin{DoxyCompactItemize}
\item 
\hyperlink{struct_validate4_1_1_retrieve_pre_abd6dc10b410baedef42cc2261ca8f143}{Retrieve\+Pre} (\hyperlink{class_validate4_a49c80b3d101be19542a4341c2387603a}{alphabet\+\_\+vector\+\_\+type} $\ast$\+\_\+t, \hyperlink{class_validate4_a46ea31a0a4b23f583806792160421d15}{size\+\_\+vector\+\_\+type} $\ast$\+\_\+sa, \hyperlink{class_validate4_aff3135d55f749682df7e0f97c82a6769}{pair3\+\_\+less\+\_\+sorter\+\_\+1st\+\_\+type} $\ast$\&\+\_\+pre\+\_\+item\+\_\+of\+\_\+lms\+\_\+sorter, \hyperlink{class_validate4_a4c1bcea817e34f96d95c99e0694bbfc8}{triple2\+\_\+less\+\_\+sorter\+\_\+1st\+\_\+type} $\ast$\&\+\_\+pre\+\_\+item\+\_\+of\+\_\+l\+\_\+sorter, \hyperlink{class_validate4_aa35a0ebf404821377118043d536be262}{triple2\+\_\+great\+\_\+sorter\+\_\+1st\+\_\+type} $\ast$\&\+\_\+pre\+\_\+item\+\_\+of\+\_\+s\+\_\+sorter, \hyperlink{struct_validate4_1_1_bkt_info}{Bkt\+Info} \&\+\_\+bkt\+\_\+info)
\begin{DoxyCompactList}\small\item\em ctor \end{DoxyCompactList}\end{DoxyCompactItemize}

\subsection{Detailed Description}
\subsubsection*{template$<$typename alphabet\+\_\+type, typename alphabet\+\_\+extension\+\_\+type, typename size\+\_\+type$>$\newline
struct Validate4$<$ alphabet\+\_\+type, alphabet\+\_\+extension\+\_\+type, size\+\_\+type $>$\+::\+Retrieve\+Pre}

retrieve preceding items (pre\+\_\+ch \& pre\+\_\+t) for L-\/type, S-\/type and L\+MS suffixes

\subsection{Constructor \& Destructor Documentation}
\mbox{\Hypertarget{struct_validate4_1_1_retrieve_pre_abd6dc10b410baedef42cc2261ca8f143}\label{struct_validate4_1_1_retrieve_pre_abd6dc10b410baedef42cc2261ca8f143}} 
\index{Validate4\+::\+Retrieve\+Pre@{Validate4\+::\+Retrieve\+Pre}!Retrieve\+Pre@{Retrieve\+Pre}}
\index{Retrieve\+Pre@{Retrieve\+Pre}!Validate4\+::\+Retrieve\+Pre@{Validate4\+::\+Retrieve\+Pre}}
\subsubsection{\texorpdfstring{Retrieve\+Pre()}{RetrievePre()}}
{\footnotesize\ttfamily template$<$typename alphabet\+\_\+type , typename alphabet\+\_\+extension\+\_\+type , typename size\+\_\+type $>$ \\
\hyperlink{class_validate4}{Validate4}$<$ alphabet\+\_\+type, alphabet\+\_\+extension\+\_\+type, size\+\_\+type $>$\+::Retrieve\+Pre\+::\+Retrieve\+Pre (\begin{DoxyParamCaption}\item[{\hyperlink{class_validate4_a49c80b3d101be19542a4341c2387603a}{alphabet\+\_\+vector\+\_\+type} $\ast$}]{\+\_\+t,  }\item[{\hyperlink{class_validate4_a46ea31a0a4b23f583806792160421d15}{size\+\_\+vector\+\_\+type} $\ast$}]{\+\_\+sa,  }\item[{\hyperlink{class_validate4_aff3135d55f749682df7e0f97c82a6769}{pair3\+\_\+less\+\_\+sorter\+\_\+1st\+\_\+type} $\ast$\&}]{\+\_\+pre\+\_\+item\+\_\+of\+\_\+lms\+\_\+sorter,  }\item[{\hyperlink{class_validate4_a4c1bcea817e34f96d95c99e0694bbfc8}{triple2\+\_\+less\+\_\+sorter\+\_\+1st\+\_\+type} $\ast$\&}]{\+\_\+pre\+\_\+item\+\_\+of\+\_\+l\+\_\+sorter,  }\item[{\hyperlink{class_validate4_aa35a0ebf404821377118043d536be262}{triple2\+\_\+great\+\_\+sorter\+\_\+1st\+\_\+type} $\ast$\&}]{\+\_\+pre\+\_\+item\+\_\+of\+\_\+s\+\_\+sorter,  }\item[{\hyperlink{struct_validate4_1_1_bkt_info}{Bkt\+Info} \&}]{\+\_\+bkt\+\_\+info }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

ctor

The documentation for this struct was generated from the following file\+:\begin{DoxyCompactItemize}
\item 
code/src/\hyperlink{validate4_8h}{validate4.\+h}\end{DoxyCompactItemize}