### 2017-01-24 archive/latex/structchoose__int__types.tex

\hypertarget{structchoose__int__types}{}\section{choose\+\_\+int\+\_\+types$<$ PS $>$ Struct Template Reference}
\label{structchoose__int__types}\index{choose\+\_\+int\+\_\+types$<$ P\+S $>$@{choose\+\_\+int\+\_\+types$<$ P\+S $>$}}

template for choosing int/uint type according to pointer size

{\ttfamily \#include $<$types.\+h$>$}

\subsection{Detailed Description}
\subsubsection*{template$<$int PS$>$\newline
struct choose\+\_\+int\+\_\+types$<$ P\+S $>$}

template for choosing int/uint type according to pointer size

The documentation for this struct was generated from the following file\+:\begin{DoxyCompactItemize}
\item 
code/src/common/\hyperlink{types_8h}{types.\+h}\end{DoxyCompactItemize}