### 2017-01-24 archive/latex/structchoose__int__types_3_014_01_4.tex

\hypertarget{structchoose__int__types_3_014_01_4}{}\section{choose\+\_\+int\+\_\+types$<$ 4 $>$ Struct Template Reference}
\label{structchoose__int__types_3_014_01_4}\index{choose\+\_\+int\+\_\+types$<$ 4 $>$@{choose\+\_\+int\+\_\+types$<$ 4 $>$}}

PS = 4, int32/uint32.

{\ttfamily \#include $<$types.\+h$>$}

\subsection*{Public Types}
\begin{DoxyCompactItemize}
\item 
using \hyperlink{structchoose__int__types_3_014_01_4_af6cc8894e99e97c539077c9e640ed478}{int\+\_\+type} = \hyperlink{types_8h_a3d594eb72953c94a18a03d929ebd9167}{int32}
\item 
using \hyperlink{structchoose__int__types_3_014_01_4_a3c2319206150de67a68544e265dabb29}{uint\+\_\+type} = \hyperlink{types_8h_a8f95e75e58492e87412191fabadd8ca8}{uint32}
\end{DoxyCompactItemize}

\subsection{Detailed Description}
\subsubsection*{template$<$$>$\newline
struct choose\+\_\+int\+\_\+types$<$ 4 $>$}

PS = 4, int32/uint32.

32-\/bit operating system, support 32-\/bit address space

\subsection{Member Typedef Documentation}
\mbox{\Hypertarget{structchoose__int__types_3_014_01_4_af6cc8894e99e97c539077c9e640ed478}\label{structchoose__int__types_3_014_01_4_af6cc8894e99e97c539077c9e640ed478}} 
\index{choose\+\_\+int\+\_\+types$<$ 4 $>$@{choose\+\_\+int\+\_\+types$<$ 4 $>$}!int\+\_\+type@{int\+\_\+type}}
\index{int\+\_\+type@{int\+\_\+type}!choose\+\_\+int\+\_\+types$<$ 4 $>$@{choose\+\_\+int\+\_\+types$<$ 4 $>$}}
\subsubsection{\texorpdfstring{int\+\_\+type}{int\_type}}
{\footnotesize\ttfamily using \hyperlink{structchoose__int__types}{choose\+\_\+int\+\_\+types}$<$ 4 $>$\+::\hyperlink{structchoose__int__types_3_014_01_4_af6cc8894e99e97c539077c9e640ed478}{int\+\_\+type} =  \hyperlink{types_8h_a3d594eb72953c94a18a03d929ebd9167}{int32}}

\mbox{\Hypertarget{structchoose__int__types_3_014_01_4_a3c2319206150de67a68544e265dabb29}\label{structchoose__int__types_3_014_01_4_a3c2319206150de67a68544e265dabb29}} 
\index{choose\+\_\+int\+\_\+types$<$ 4 $>$@{choose\+\_\+int\+\_\+types$<$ 4 $>$}!uint\+\_\+type@{uint\+\_\+type}}
\index{uint\+\_\+type@{uint\+\_\+type}!choose\+\_\+int\+\_\+types$<$ 4 $>$@{choose\+\_\+int\+\_\+types$<$ 4 $>$}}
\subsubsection{\texorpdfstring{uint\+\_\+type}{uint\_type}}
{\footnotesize\ttfamily using \hyperlink{structchoose__int__types}{choose\+\_\+int\+\_\+types}$<$ 4 $>$\+::\hyperlink{structchoose__int__types_3_014_01_4_a3c2319206150de67a68544e265dabb29}{uint\+\_\+type} =  \hyperlink{types_8h_a8f95e75e58492e87412191fabadd8ca8}{uint32}}

The documentation for this struct was generated from the following file\+:\begin{DoxyCompactItemize}
\item 
code/src/common/\hyperlink{types_8h}{types.\+h}\end{DoxyCompactItemize}