### 2017-01-24 archive/latex/structchoose__int__types_3_018_01_4.tex

\hypertarget{structchoose__int__types_3_018_01_4}{}\section{choose\+\_\+int\+\_\+types$<$ 8 $>$ Struct Template Reference}
\label{structchoose__int__types_3_018_01_4}\index{choose\+\_\+int\+\_\+types$<$ 8 $>$@{choose\+\_\+int\+\_\+types$<$ 8 $>$}}

pointer size = 8, 64-\/bit

{\ttfamily \#include $<$types.\+h$>$}

\subsection*{Public Types}
\begin{DoxyCompactItemize}
\item 
using \hyperlink{structchoose__int__types_3_018_01_4_a4dde5a87f22a82a381bbae5f6d205d05}{int\+\_\+type} = \hyperlink{types_8h_a552928ab323811c9694f5b7c9f53d0fb}{int64}
\item 
using \hyperlink{structchoose__int__types_3_018_01_4_a4e70f94b235eba865f95080e668f666a}{uint\+\_\+type} = \hyperlink{types_8h_a60e8696a4678cd348e991a1f172e53f7}{uint64}
\end{DoxyCompactItemize}

\subsection{Detailed Description}
\subsubsection*{template$<$$>$\newline
struct choose\+\_\+int\+\_\+types$<$ 8 $>$}

pointer size = 8, 64-\/bit

64-\/bit operating system, support 64-\/bit address space

\subsection{Member Typedef Documentation}
\mbox{\Hypertarget{structchoose__int__types_3_018_01_4_a4dde5a87f22a82a381bbae5f6d205d05}\label{structchoose__int__types_3_018_01_4_a4dde5a87f22a82a381bbae5f6d205d05}} 
\index{choose\+\_\+int\+\_\+types$<$ 8 $>$@{choose\+\_\+int\+\_\+types$<$ 8 $>$}!int\+\_\+type@{int\+\_\+type}}
\index{int\+\_\+type@{int\+\_\+type}!choose\+\_\+int\+\_\+types$<$ 8 $>$@{choose\+\_\+int\+\_\+types$<$ 8 $>$}}
\subsubsection{\texorpdfstring{int\+\_\+type}{int\_type}}
{\footnotesize\ttfamily using \hyperlink{structchoose__int__types}{choose\+\_\+int\+\_\+types}$<$ 8 $>$\+::\hyperlink{structchoose__int__types_3_018_01_4_a4dde5a87f22a82a381bbae5f6d205d05}{int\+\_\+type} =  \hyperlink{types_8h_a552928ab323811c9694f5b7c9f53d0fb}{int64}}

\mbox{\Hypertarget{structchoose__int__types_3_018_01_4_a4e70f94b235eba865f95080e668f666a}\label{structchoose__int__types_3_018_01_4_a4e70f94b235eba865f95080e668f666a}} 
\index{choose\+\_\+int\+\_\+types$<$ 8 $>$@{choose\+\_\+int\+\_\+types$<$ 8 $>$}!uint\+\_\+type@{uint\+\_\+type}}
\index{uint\+\_\+type@{uint\+\_\+type}!choose\+\_\+int\+\_\+types$<$ 8 $>$@{choose\+\_\+int\+\_\+types$<$ 8 $>$}}
\subsubsection{\texorpdfstring{uint\+\_\+type}{uint\_type}}
{\footnotesize\ttfamily using \hyperlink{structchoose__int__types}{choose\+\_\+int\+\_\+types}$<$ 8 $>$\+::\hyperlink{structchoose__int__types_3_018_01_4_a4e70f94b235eba865f95080e668f666a}{uint\+\_\+type} =  \hyperlink{types_8h_a60e8696a4678cd348e991a1f172e53f7}{uint64}}

The documentation for this struct was generated from the following file\+:\begin{DoxyCompactItemize}
\item 
code/src/common/\hyperlink{types_8h}{types.\+h}\end{DoxyCompactItemize}