### 2017-01-24 archive/latex/structpair.tex

\hypertarget{structpair}{}\section{pair$<$ T1, T2 $>$ Struct Template Reference}
\label{structpair}\index{pair$<$ T1, T2 $>$@{pair$<$ T1, T2 $>$}}

structure pair

{\ttfamily \#include $<$tuples.\+h$>$}

\subsection*{Public Member Functions}
\begin{DoxyCompactItemize}
\item 
\hyperlink{structpair_a1dba17bb5c888d7ed8ce7b7c4acc52de}{pair} ()
\begin{DoxyCompactList}\small\item\em constructor, default \end{DoxyCompactList}\item 
\hyperlink{structpair_ab75f77e9a1f726301f66d3be6b1e184a}{pair} (const \hyperlink{structpair}{pair} \&\+\_\+item)
\begin{DoxyCompactList}\small\item\em constructor, copy \end{DoxyCompactList}\item 
\hyperlink{structpair_a5674a7a5327419692abe6160ae202859}{pair} (const T1 \&\+\_\+first, const T2 \&\+\_\+second)
\begin{DoxyCompactList}\small\item\em constructor, component \end{DoxyCompactList}\end{DoxyCompactItemize}
\subsection*{Static Public Member Functions}
\begin{DoxyCompactItemize}
\item 
static \hyperlink{structpair}{pair} \& \hyperlink{structpair_abca1ec29c5040ce16bd3993e80d42923}{min\+\_\+value} ()
\begin{DoxyCompactList}\small\item\em min value \end{DoxyCompactList}\item 
static \hyperlink{structpair}{pair} \& \hyperlink{structpair_a89b59c46e545a619e96c7bd5d66a8ef6}{max\+\_\+value} ()
\begin{DoxyCompactList}\small\item\em max vlaue \end{DoxyCompactList}\end{DoxyCompactItemize}
\subsection*{Public Attributes}
\begin{DoxyCompactItemize}
\item 
T1 \hyperlink{structpair_aed925258783c91efbc5432030709d54e}{first}
\item 
T2 \hyperlink{structpair_a63ae1bbd5a6ce6bd6167b5359fa7a01f}{second}
\end{DoxyCompactItemize}
\subsection*{Friends}
\begin{DoxyCompactItemize}
\item 
std\+::ostream \& \hyperlink{structpair_a9cd7ed0b2b2d0a6d5977137d9cf35a0c}{operator$<$$<$} (std\+::ostream \+\_\+os, const \hyperlink{structpair}{pair} \&\+\_\+item)
\begin{DoxyCompactList}\small\item\em output \end{DoxyCompactList}\end{DoxyCompactItemize}

\subsection{Detailed Description}
\subsubsection*{template$<$typename T1, typename T2$>$\newline
struct pair$<$ T1, T2 $>$}

structure pair

\subsection{Constructor \& Destructor Documentation}
\mbox{\Hypertarget{structpair_a1dba17bb5c888d7ed8ce7b7c4acc52de}\label{structpair_a1dba17bb5c888d7ed8ce7b7c4acc52de}} 
\index{pair@{pair}!pair@{pair}}
\index{pair@{pair}!pair@{pair}}
\subsubsection{\texorpdfstring{pair()}{pair()}\hspace{0.1cm}{\footnotesize\ttfamily [1/3]}}
{\footnotesize\ttfamily template$<$typename T1 , typename T2 $>$ \\
\hyperlink{structpair}{pair}$<$ T1, T2 $>$\+::\hyperlink{structpair}{pair} (\begin{DoxyParamCaption}{ }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

constructor, default

\mbox{\Hypertarget{structpair_ab75f77e9a1f726301f66d3be6b1e184a}\label{structpair_ab75f77e9a1f726301f66d3be6b1e184a}} 
\index{pair@{pair}!pair@{pair}}
\index{pair@{pair}!pair@{pair}}
\subsubsection{\texorpdfstring{pair()}{pair()}\hspace{0.1cm}{\footnotesize\ttfamily [2/3]}}
{\footnotesize\ttfamily template$<$typename T1 , typename T2 $>$ \\
\hyperlink{structpair}{pair}$<$ T1, T2 $>$\+::\hyperlink{structpair}{pair} (\begin{DoxyParamCaption}\item[{const \hyperlink{structpair}{pair}$<$ T1, T2 $>$ \&}]{\+\_\+item }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

constructor, copy

\mbox{\Hypertarget{structpair_a5674a7a5327419692abe6160ae202859}\label{structpair_a5674a7a5327419692abe6160ae202859}} 
\index{pair@{pair}!pair@{pair}}
\index{pair@{pair}!pair@{pair}}
\subsubsection{\texorpdfstring{pair()}{pair()}\hspace{0.1cm}{\footnotesize\ttfamily [3/3]}}
{\footnotesize\ttfamily template$<$typename T1 , typename T2 $>$ \\
\hyperlink{structpair}{pair}$<$ T1, T2 $>$\+::\hyperlink{structpair}{pair} (\begin{DoxyParamCaption}\item[{const T1 \&}]{\+\_\+first,  }\item[{const T2 \&}]{\+\_\+second }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

constructor, component

\subsection{Member Function Documentation}
\mbox{\Hypertarget{structpair_a89b59c46e545a619e96c7bd5d66a8ef6}\label{structpair_a89b59c46e545a619e96c7bd5d66a8ef6}} 
\index{pair@{pair}!max\+\_\+value@{max\+\_\+value}}
\index{max\+\_\+value@{max\+\_\+value}!pair@{pair}}
\subsubsection{\texorpdfstring{max\+\_\+value()}{max\_value()}}
{\footnotesize\ttfamily template$<$typename T1 , typename T2 $>$ \\
static \hyperlink{structpair}{pair}\& \hyperlink{structpair}{pair}$<$ T1, T2 $>$\+::max\+\_\+value (\begin{DoxyParamCaption}{ }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}, {\ttfamily [static]}}

max vlaue

\mbox{\Hypertarget{structpair_abca1ec29c5040ce16bd3993e80d42923}\label{structpair_abca1ec29c5040ce16bd3993e80d42923}} 
\index{pair@{pair}!min\+\_\+value@{min\+\_\+value}}
\index{min\+\_\+value@{min\+\_\+value}!pair@{pair}}
\subsubsection{\texorpdfstring{min\+\_\+value()}{min\_value()}}
{\footnotesize\ttfamily template$<$typename T1 , typename T2 $>$ \\
static \hyperlink{structpair}{pair}\& \hyperlink{structpair}{pair}$<$ T1, T2 $>$\+::min\+\_\+value (\begin{DoxyParamCaption}{ }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}, {\ttfamily [static]}}

min value

\subsection{Friends And Related Function Documentation}
\mbox{\Hypertarget{structpair_a9cd7ed0b2b2d0a6d5977137d9cf35a0c}\label{structpair_a9cd7ed0b2b2d0a6d5977137d9cf35a0c}} 
\index{pair@{pair}!operator$<$$<$@{operator$<$$<$}}
\index{operator$<$$<$@{operator$<$$<$}!pair@{pair}}
\subsubsection{\texorpdfstring{operator$<$$<$}{operator<<}}
{\footnotesize\ttfamily template$<$typename T1 , typename T2 $>$ \\
std\+::ostream\& operator$<$$<$ (\begin{DoxyParamCaption}\item[{std\+::ostream}]{\+\_\+os,  }\item[{const \hyperlink{structpair}{pair}$<$ T1, T2 $>$ \&}]{\+\_\+item }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [friend]}}

output

\subsection{Member Data Documentation}
\mbox{\Hypertarget{structpair_aed925258783c91efbc5432030709d54e}\label{structpair_aed925258783c91efbc5432030709d54e}} 
\index{pair@{pair}!first@{first}}
\index{first@{first}!pair@{pair}}
\subsubsection{\texorpdfstring{first}{first}}
{\footnotesize\ttfamily template$<$typename T1 , typename T2 $>$ \\
T1 \hyperlink{structpair}{pair}$<$ T1, T2 $>$\+::first}

\mbox{\Hypertarget{structpair_a63ae1bbd5a6ce6bd6167b5359fa7a01f}\label{structpair_a63ae1bbd5a6ce6bd6167b5359fa7a01f}} 
\index{pair@{pair}!second@{second}}
\index{second@{second}!pair@{pair}}
\subsubsection{\texorpdfstring{second}{second}}
{\footnotesize\ttfamily template$<$typename T1 , typename T2 $>$ \\
T2 \hyperlink{structpair}{pair}$<$ T1, T2 $>$\+::second}

The documentation for this struct was generated from the following file\+:\begin{DoxyCompactItemize}
\item 
code/src/common/\hyperlink{tuples_8h}{tuples.\+h}\end{DoxyCompactItemize}