### 2017-01-24 archive/latex/structquadruple.tex

\hypertarget{structquadruple}{}\section{quadruple$<$ T1, T2, T3, T4 $>$ Struct Template Reference}
\label{structquadruple}\index{quadruple$<$ T1, T2, T3, T4 $>$@{quadruple$<$ T1, T2, T3, T4 $>$}}

structure quadruple

{\ttfamily \#include $<$tuples.\+h$>$}

\subsection*{Public Member Functions}
\begin{DoxyCompactItemize}
\item 
\hyperlink{structquadruple_ab4b318f9f20c5f861c0b7445bf806da8}{quadruple} ()
\begin{DoxyCompactList}\small\item\em constructor, default \end{DoxyCompactList}\item 
\hyperlink{structquadruple_a24c753a7c00f8325bf7f738815e89d06}{quadruple} (const \hyperlink{structquadruple}{quadruple} \&\+\_\+item)
\begin{DoxyCompactList}\small\item\em constructor, copy \end{DoxyCompactList}\item 
\hyperlink{structquadruple_a23034a914f2841aa59c833bb3923f1b6}{quadruple} (const T1 \&\+\_\+first, const T2 \&\+\_\+second, const T3 \&\+\_\+third, const T4 \&\+\_\+forth)
\begin{DoxyCompactList}\small\item\em constructor, component \end{DoxyCompactList}\end{DoxyCompactItemize}
\subsection*{Static Public Member Functions}
\begin{DoxyCompactItemize}
\item 
static \hyperlink{structquadruple}{quadruple} \& \hyperlink{structquadruple_a97a2767881ac4cfb52513827b4f51588}{min\+\_\+value} ()
\begin{DoxyCompactList}\small\item\em min value \end{DoxyCompactList}\item 
static \hyperlink{structquadruple}{quadruple} \& \hyperlink{structquadruple_a1361dd9869f7323e5fc33ab07864e128}{max\+\_\+value} ()
\begin{DoxyCompactList}\small\item\em max value \end{DoxyCompactList}\end{DoxyCompactItemize}
\subsection*{Public Attributes}
\begin{DoxyCompactItemize}
\item 
T1 \hyperlink{structquadruple_a45642eeda831800cd5c540037822ac65}{first}
\item 
T2 \hyperlink{structquadruple_aa93ca1ef60012d07903333d7b16bb32b}{second}
\item 
T3 \hyperlink{structquadruple_af1ada0ce5f85bbc7df2c5bfb113b442d}{third}
\item 
T4 \hyperlink{structquadruple_abab1a25fd1ed39d2605eaa9371de04ca}{forth}
\end{DoxyCompactItemize}
\subsection*{Friends}
\begin{DoxyCompactItemize}
\item 
std\+::ostream \& \hyperlink{structquadruple_af619be80f82068b7b05a8d943196f024}{operator$<$$<$} (std\+::ostream \+\_\+os, const \hyperlink{structquadruple}{quadruple} \&\+\_\+item)
\begin{DoxyCompactList}\small\item\em output \end{DoxyCompactList}\end{DoxyCompactItemize}

\subsection{Detailed Description}
\subsubsection*{template$<$typename T1, typename T2, typename T3, typename T4$>$\newline
struct quadruple$<$ T1, T2, T3, T4 $>$}

structure quadruple

\subsection{Constructor \& Destructor Documentation}
\mbox{\Hypertarget{structquadruple_ab4b318f9f20c5f861c0b7445bf806da8}\label{structquadruple_ab4b318f9f20c5f861c0b7445bf806da8}} 
\index{quadruple@{quadruple}!quadruple@{quadruple}}
\index{quadruple@{quadruple}!quadruple@{quadruple}}
\subsubsection{\texorpdfstring{quadruple()}{quadruple()}\hspace{0.1cm}{\footnotesize\ttfamily [1/3]}}
{\footnotesize\ttfamily template$<$typename T1 , typename T2 , typename T3 , typename T4 $>$ \\
\hyperlink{structquadruple}{quadruple}$<$ T1, T2, T3, T4 $>$\+::\hyperlink{structquadruple}{quadruple} (\begin{DoxyParamCaption}{ }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

constructor, default

\mbox{\Hypertarget{structquadruple_a24c753a7c00f8325bf7f738815e89d06}\label{structquadruple_a24c753a7c00f8325bf7f738815e89d06}} 
\index{quadruple@{quadruple}!quadruple@{quadruple}}
\index{quadruple@{quadruple}!quadruple@{quadruple}}
\subsubsection{\texorpdfstring{quadruple()}{quadruple()}\hspace{0.1cm}{\footnotesize\ttfamily [2/3]}}
{\footnotesize\ttfamily template$<$typename T1 , typename T2 , typename T3 , typename T4 $>$ \\
\hyperlink{structquadruple}{quadruple}$<$ T1, T2, T3, T4 $>$\+::\hyperlink{structquadruple}{quadruple} (\begin{DoxyParamCaption}\item[{const \hyperlink{structquadruple}{quadruple}$<$ T1, T2, T3, T4 $>$ \&}]{\+\_\+item }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

constructor, copy

\mbox{\Hypertarget{structquadruple_a23034a914f2841aa59c833bb3923f1b6}\label{structquadruple_a23034a914f2841aa59c833bb3923f1b6}} 
\index{quadruple@{quadruple}!quadruple@{quadruple}}
\index{quadruple@{quadruple}!quadruple@{quadruple}}
\subsubsection{\texorpdfstring{quadruple()}{quadruple()}\hspace{0.1cm}{\footnotesize\ttfamily [3/3]}}
{\footnotesize\ttfamily template$<$typename T1 , typename T2 , typename T3 , typename T4 $>$ \\
\hyperlink{structquadruple}{quadruple}$<$ T1, T2, T3, T4 $>$\+::\hyperlink{structquadruple}{quadruple} (\begin{DoxyParamCaption}\item[{const T1 \&}]{\+\_\+first,  }\item[{const T2 \&}]{\+\_\+second,  }\item[{const T3 \&}]{\+\_\+third,  }\item[{const T4 \&}]{\+\_\+forth }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

constructor, component

\subsection{Member Function Documentation}
\mbox{\Hypertarget{structquadruple_a1361dd9869f7323e5fc33ab07864e128}\label{structquadruple_a1361dd9869f7323e5fc33ab07864e128}} 
\index{quadruple@{quadruple}!max\+\_\+value@{max\+\_\+value}}
\index{max\+\_\+value@{max\+\_\+value}!quadruple@{quadruple}}
\subsubsection{\texorpdfstring{max\+\_\+value()}{max\_value()}}
{\footnotesize\ttfamily template$<$typename T1 , typename T2 , typename T3 , typename T4 $>$ \\
static \hyperlink{structquadruple}{quadruple}\& \hyperlink{structquadruple}{quadruple}$<$ T1, T2, T3, T4 $>$\+::max\+\_\+value (\begin{DoxyParamCaption}{ }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}, {\ttfamily [static]}}

max value

\mbox{\Hypertarget{structquadruple_a97a2767881ac4cfb52513827b4f51588}\label{structquadruple_a97a2767881ac4cfb52513827b4f51588}} 
\index{quadruple@{quadruple}!min\+\_\+value@{min\+\_\+value}}
\index{min\+\_\+value@{min\+\_\+value}!quadruple@{quadruple}}
\subsubsection{\texorpdfstring{min\+\_\+value()}{min\_value()}}
{\footnotesize\ttfamily template$<$typename T1 , typename T2 , typename T3 , typename T4 $>$ \\
static \hyperlink{structquadruple}{quadruple}\& \hyperlink{structquadruple}{quadruple}$<$ T1, T2, T3, T4 $>$\+::min\+\_\+value (\begin{DoxyParamCaption}{ }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}, {\ttfamily [static]}}

min value

\subsection{Friends And Related Function Documentation}
\mbox{\Hypertarget{structquadruple_af619be80f82068b7b05a8d943196f024}\label{structquadruple_af619be80f82068b7b05a8d943196f024}} 
\index{quadruple@{quadruple}!operator$<$$<$@{operator$<$$<$}}
\index{operator$<$$<$@{operator$<$$<$}!quadruple@{quadruple}}
\subsubsection{\texorpdfstring{operator$<$$<$}{operator<<}}
{\footnotesize\ttfamily template$<$typename T1 , typename T2 , typename T3 , typename T4 $>$ \\
std\+::ostream\& operator$<$$<$ (\begin{DoxyParamCaption}\item[{std\+::ostream}]{\+\_\+os,  }\item[{const \hyperlink{structquadruple}{quadruple}$<$ T1, T2, T3, T4 $>$ \&}]{\+\_\+item }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [friend]}}

output

\subsection{Member Data Documentation}
\mbox{\Hypertarget{structquadruple_a45642eeda831800cd5c540037822ac65}\label{structquadruple_a45642eeda831800cd5c540037822ac65}} 
\index{quadruple@{quadruple}!first@{first}}
\index{first@{first}!quadruple@{quadruple}}
\subsubsection{\texorpdfstring{first}{first}}
{\footnotesize\ttfamily template$<$typename T1 , typename T2 , typename T3 , typename T4 $>$ \\
T1 \hyperlink{structquadruple}{quadruple}$<$ T1, T2, T3, T4 $>$\+::first}

\mbox{\Hypertarget{structquadruple_abab1a25fd1ed39d2605eaa9371de04ca}\label{structquadruple_abab1a25fd1ed39d2605eaa9371de04ca}} 
\index{quadruple@{quadruple}!forth@{forth}}
\index{forth@{forth}!quadruple@{quadruple}}
\subsubsection{\texorpdfstring{forth}{forth}}
{\footnotesize\ttfamily template$<$typename T1 , typename T2 , typename T3 , typename T4 $>$ \\
T4 \hyperlink{structquadruple}{quadruple}$<$ T1, T2, T3, T4 $>$\+::forth}

\mbox{\Hypertarget{structquadruple_aa93ca1ef60012d07903333d7b16bb32b}\label{structquadruple_aa93ca1ef60012d07903333d7b16bb32b}} 
\index{quadruple@{quadruple}!second@{second}}
\index{second@{second}!quadruple@{quadruple}}
\subsubsection{\texorpdfstring{second}{second}}
{\footnotesize\ttfamily template$<$typename T1 , typename T2 , typename T3 , typename T4 $>$ \\
T2 \hyperlink{structquadruple}{quadruple}$<$ T1, T2, T3, T4 $>$\+::second}

\mbox{\Hypertarget{structquadruple_af1ada0ce5f85bbc7df2c5bfb113b442d}\label{structquadruple_af1ada0ce5f85bbc7df2c5bfb113b442d}} 
\index{quadruple@{quadruple}!third@{third}}
\index{third@{third}!quadruple@{quadruple}}
\subsubsection{\texorpdfstring{third}{third}}
{\footnotesize\ttfamily template$<$typename T1 , typename T2 , typename T3 , typename T4 $>$ \\
T3 \hyperlink{structquadruple}{quadruple}$<$ T1, T2, T3, T4 $>$\+::third}

The documentation for this struct was generated from the following file\+:\begin{DoxyCompactItemize}
\item 
code/src/common/\hyperlink{tuples_8h}{tuples.\+h}\end{DoxyCompactItemize}