### 2017-01-24 archive/latex/structtriple.tex

\hypertarget{structtriple}{}\section{triple$<$ T1, T2, T3 $>$ Struct Template Reference}
\label{structtriple}\index{triple$<$ T1, T2, T3 $>$@{triple$<$ T1, T2, T3 $>$}}

structure triple

{\ttfamily \#include $<$tuples.\+h$>$}

\subsection*{Public Member Functions}
\begin{DoxyCompactItemize}
\item 
\hyperlink{structtriple_a3bdf8136df773fff422b4cbb354cf749}{triple} ()
\begin{DoxyCompactList}\small\item\em constructor, default \end{DoxyCompactList}\item 
\hyperlink{structtriple_a88688d689a93bf4b8a25bc29728a53b5}{triple} (const \hyperlink{structtriple}{triple} \&\+\_\+item)
\begin{DoxyCompactList}\small\item\em constructor, copy \end{DoxyCompactList}\item 
\hyperlink{structtriple_ac34fdb8619fdc11fb442e2058d35d840}{triple} (const T1 \&\+\_\+first, const T2 \&\+\_\+second, const T3 \&\+\_\+third)
\begin{DoxyCompactList}\small\item\em constructor, component \end{DoxyCompactList}\end{DoxyCompactItemize}
\subsection*{Static Public Member Functions}
\begin{DoxyCompactItemize}
\item 
static \hyperlink{structtriple}{triple} \& \hyperlink{structtriple_a91d8338063226b63eea74e56377a162b}{min\+\_\+value} ()
\begin{DoxyCompactList}\small\item\em min value \end{DoxyCompactList}\item 
static \hyperlink{structtriple}{triple} \& \hyperlink{structtriple_a73d954bee19bf6493205e794a6448e8f}{max\+\_\+value} ()
\begin{DoxyCompactList}\small\item\em max value \end{DoxyCompactList}\end{DoxyCompactItemize}
\subsection*{Public Attributes}
\begin{DoxyCompactItemize}
\item 
T1 \hyperlink{structtriple_a6aa8508e6e6f859dda04f144f361c0d2}{first}
\item 
T2 \hyperlink{structtriple_a41050194c8c7984707c7ad13c96e8631}{second}
\item 
T3 \hyperlink{structtriple_a5675f08ea7996033edbcf7074404cdd8}{third}
\end{DoxyCompactItemize}
\subsection*{Friends}
\begin{DoxyCompactItemize}
\item 
std\+::ostream \& \hyperlink{structtriple_a213e2d21647f20d6287690438970f212}{operator$<$$<$} (std\+::ostream \+\_\+os, const \hyperlink{structtriple}{triple} \&\+\_\+item)
\begin{DoxyCompactList}\small\item\em output \end{DoxyCompactList}\end{DoxyCompactItemize}

\subsection{Detailed Description}
\subsubsection*{template$<$typename T1, typename T2, typename T3$>$\newline
struct triple$<$ T1, T2, T3 $>$}

structure triple

\subsection{Constructor \& Destructor Documentation}
\mbox{\Hypertarget{structtriple_a3bdf8136df773fff422b4cbb354cf749}\label{structtriple_a3bdf8136df773fff422b4cbb354cf749}} 
\index{triple@{triple}!triple@{triple}}
\index{triple@{triple}!triple@{triple}}
\subsubsection{\texorpdfstring{triple()}{triple()}\hspace{0.1cm}{\footnotesize\ttfamily [1/3]}}
{\footnotesize\ttfamily template$<$typename T1 , typename T2 , typename T3 $>$ \\
\hyperlink{structtriple}{triple}$<$ T1, T2, T3 $>$\+::\hyperlink{structtriple}{triple} (\begin{DoxyParamCaption}{ }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

constructor, default

\mbox{\Hypertarget{structtriple_a88688d689a93bf4b8a25bc29728a53b5}\label{structtriple_a88688d689a93bf4b8a25bc29728a53b5}} 
\index{triple@{triple}!triple@{triple}}
\index{triple@{triple}!triple@{triple}}
\subsubsection{\texorpdfstring{triple()}{triple()}\hspace{0.1cm}{\footnotesize\ttfamily [2/3]}}
{\footnotesize\ttfamily template$<$typename T1 , typename T2 , typename T3 $>$ \\
\hyperlink{structtriple}{triple}$<$ T1, T2, T3 $>$\+::\hyperlink{structtriple}{triple} (\begin{DoxyParamCaption}\item[{const \hyperlink{structtriple}{triple}$<$ T1, T2, T3 $>$ \&}]{\+\_\+item }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

constructor, copy

\mbox{\Hypertarget{structtriple_ac34fdb8619fdc11fb442e2058d35d840}\label{structtriple_ac34fdb8619fdc11fb442e2058d35d840}} 
\index{triple@{triple}!triple@{triple}}
\index{triple@{triple}!triple@{triple}}
\subsubsection{\texorpdfstring{triple()}{triple()}\hspace{0.1cm}{\footnotesize\ttfamily [3/3]}}
{\footnotesize\ttfamily template$<$typename T1 , typename T2 , typename T3 $>$ \\
\hyperlink{structtriple}{triple}$<$ T1, T2, T3 $>$\+::\hyperlink{structtriple}{triple} (\begin{DoxyParamCaption}\item[{const T1 \&}]{\+\_\+first,  }\item[{const T2 \&}]{\+\_\+second,  }\item[{const T3 \&}]{\+\_\+third }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}}

constructor, component

\subsection{Member Function Documentation}
\mbox{\Hypertarget{structtriple_a73d954bee19bf6493205e794a6448e8f}\label{structtriple_a73d954bee19bf6493205e794a6448e8f}} 
\index{triple@{triple}!max\+\_\+value@{max\+\_\+value}}
\index{max\+\_\+value@{max\+\_\+value}!triple@{triple}}
\subsubsection{\texorpdfstring{max\+\_\+value()}{max\_value()}}
{\footnotesize\ttfamily template$<$typename T1 , typename T2 , typename T3 $>$ \\
static \hyperlink{structtriple}{triple}\& \hyperlink{structtriple}{triple}$<$ T1, T2, T3 $>$\+::max\+\_\+value (\begin{DoxyParamCaption}{ }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}, {\ttfamily [static]}}

max value

\mbox{\Hypertarget{structtriple_a91d8338063226b63eea74e56377a162b}\label{structtriple_a91d8338063226b63eea74e56377a162b}} 
\index{triple@{triple}!min\+\_\+value@{min\+\_\+value}}
\index{min\+\_\+value@{min\+\_\+value}!triple@{triple}}
\subsubsection{\texorpdfstring{min\+\_\+value()}{min\_value()}}
{\footnotesize\ttfamily template$<$typename T1 , typename T2 , typename T3 $>$ \\
static \hyperlink{structtriple}{triple}\& \hyperlink{structtriple}{triple}$<$ T1, T2, T3 $>$\+::min\+\_\+value (\begin{DoxyParamCaption}{ }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [inline]}, {\ttfamily [static]}}

min value

\subsection{Friends And Related Function Documentation}
\mbox{\Hypertarget{structtriple_a213e2d21647f20d6287690438970f212}\label{structtriple_a213e2d21647f20d6287690438970f212}} 
\index{triple@{triple}!operator$<$$<$@{operator$<$$<$}}
\index{operator$<$$<$@{operator$<$$<$}!triple@{triple}}
\subsubsection{\texorpdfstring{operator$<$$<$}{operator<<}}
{\footnotesize\ttfamily template$<$typename T1 , typename T2 , typename T3 $>$ \\
std\+::ostream\& operator$<$$<$ (\begin{DoxyParamCaption}\item[{std\+::ostream}]{\+\_\+os,  }\item[{const \hyperlink{structtriple}{triple}$<$ T1, T2, T3 $>$ \&}]{\+\_\+item }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [friend]}}

output

\subsection{Member Data Documentation}
\mbox{\Hypertarget{structtriple_a6aa8508e6e6f859dda04f144f361c0d2}\label{structtriple_a6aa8508e6e6f859dda04f144f361c0d2}} 
\index{triple@{triple}!first@{first}}
\index{first@{first}!triple@{triple}}
\subsubsection{\texorpdfstring{first}{first}}
{\footnotesize\ttfamily template$<$typename T1 , typename T2 , typename T3 $>$ \\
T1 \hyperlink{structtriple}{triple}$<$ T1, T2, T3 $>$\+::first}

\mbox{\Hypertarget{structtriple_a41050194c8c7984707c7ad13c96e8631}\label{structtriple_a41050194c8c7984707c7ad13c96e8631}} 
\index{triple@{triple}!second@{second}}
\index{second@{second}!triple@{triple}}
\subsubsection{\texorpdfstring{second}{second}}
{\footnotesize\ttfamily template$<$typename T1 , typename T2 , typename T3 $>$ \\
T2 \hyperlink{structtriple}{triple}$<$ T1, T2, T3 $>$\+::second}

\mbox{\Hypertarget{structtriple_a5675f08ea7996033edbcf7074404cdd8}\label{structtriple_a5675f08ea7996033edbcf7074404cdd8}} 
\index{triple@{triple}!third@{third}}
\index{third@{third}!triple@{triple}}
\subsubsection{\texorpdfstring{third}{third}}
{\footnotesize\ttfamily template$<$typename T1 , typename T2 , typename T3 $>$ \\
T3 \hyperlink{structtriple}{triple}$<$ T1, T2, T3 $>$\+::third}

The documentation for this struct was generated from the following file\+:\begin{DoxyCompactItemize}
\item 
code/src/common/\hyperlink{tuples_8h}{tuples.\+h}\end{DoxyCompactItemize}