### 2017-01-24 archive/latex/structtuple__less__comparator__1st.tex

\hypertarget{structtuple__less__comparator__1st}{}\section{tuple\+\_\+less\+\_\+comparator\+\_\+1st$<$ tuple\+\_\+type $>$ Struct Template Reference}
\label{structtuple__less__comparator__1st}\index{tuple\+\_\+less\+\_\+comparator\+\_\+1st$<$ tuple\+\_\+type $>$@{tuple\+\_\+less\+\_\+comparator\+\_\+1st$<$ tuple\+\_\+type $>$}}

function object for comparing tuples by their first component in ascending order

{\ttfamily \#include $<$widget.\+h$>$}

\subsection*{Public Member Functions}
\begin{DoxyCompactItemize}
\item 
bool \hyperlink{structtuple__less__comparator__1st_ad97f49eea57d8fec456d9d00a5126deb}{operator()} (const tuple\+\_\+type \&\+\_\+a, const tuple\+\_\+type \&\+\_\+b) const
\item 
tuple\+\_\+type \hyperlink{structtuple__less__comparator__1st_a1dc22790efdf3705634387a96a848d89}{min\+\_\+value} () const
\begin{DoxyCompactList}\small\item\em min value \end{DoxyCompactList}\item 
tuple\+\_\+type \hyperlink{structtuple__less__comparator__1st_ab43723d6ebd5e36767f98828e2df83a8}{max\+\_\+value} () const
\begin{DoxyCompactList}\small\item\em max value \end{DoxyCompactList}\end{DoxyCompactItemize}

\subsection{Detailed Description}
\subsubsection*{template$<$typename tuple\+\_\+type$>$\newline
struct tuple\+\_\+less\+\_\+comparator\+\_\+1st$<$ tuple\+\_\+type $>$}

function object for comparing tuples by their first component in ascending order

\subsection{Member Function Documentation}
\mbox{\Hypertarget{structtuple__less__comparator__1st_ab43723d6ebd5e36767f98828e2df83a8}\label{structtuple__less__comparator__1st_ab43723d6ebd5e36767f98828e2df83a8}} 
\index{tuple\+\_\+less\+\_\+comparator\+\_\+1st@{tuple\+\_\+less\+\_\+comparator\+\_\+1st}!max\+\_\+value@{max\+\_\+value}}
\index{max\+\_\+value@{max\+\_\+value}!tuple\+\_\+less\+\_\+comparator\+\_\+1st@{tuple\+\_\+less\+\_\+comparator\+\_\+1st}}
\subsubsection{\texorpdfstring{max\+\_\+value()}{max\_value()}}
{\footnotesize\ttfamily template$<$typename tuple\+\_\+type $>$ \\
tuple\+\_\+type \hyperlink{structtuple__less__comparator__1st}{tuple\+\_\+less\+\_\+comparator\+\_\+1st}$<$ tuple\+\_\+type $>$\+::max\+\_\+value (\begin{DoxyParamCaption}{ }\end{DoxyParamCaption}) const\hspace{0.3cm}{\ttfamily [inline]}}

max value

\mbox{\Hypertarget{structtuple__less__comparator__1st_a1dc22790efdf3705634387a96a848d89}\label{structtuple__less__comparator__1st_a1dc22790efdf3705634387a96a848d89}} 
\index{tuple\+\_\+less\+\_\+comparator\+\_\+1st@{tuple\+\_\+less\+\_\+comparator\+\_\+1st}!min\+\_\+value@{min\+\_\+value}}
\index{min\+\_\+value@{min\+\_\+value}!tuple\+\_\+less\+\_\+comparator\+\_\+1st@{tuple\+\_\+less\+\_\+comparator\+\_\+1st}}
\subsubsection{\texorpdfstring{min\+\_\+value()}{min\_value()}}
{\footnotesize\ttfamily template$<$typename tuple\+\_\+type $>$ \\
tuple\+\_\+type \hyperlink{structtuple__less__comparator__1st}{tuple\+\_\+less\+\_\+comparator\+\_\+1st}$<$ tuple\+\_\+type $>$\+::min\+\_\+value (\begin{DoxyParamCaption}{ }\end{DoxyParamCaption}) const\hspace{0.3cm}{\ttfamily [inline]}}

min value

\mbox{\Hypertarget{structtuple__less__comparator__1st_ad97f49eea57d8fec456d9d00a5126deb}\label{structtuple__less__comparator__1st_ad97f49eea57d8fec456d9d00a5126deb}} 
\index{tuple\+\_\+less\+\_\+comparator\+\_\+1st@{tuple\+\_\+less\+\_\+comparator\+\_\+1st}!operator()@{operator()}}
\index{operator()@{operator()}!tuple\+\_\+less\+\_\+comparator\+\_\+1st@{tuple\+\_\+less\+\_\+comparator\+\_\+1st}}
\subsubsection{\texorpdfstring{operator()()}{operator()()}}
{\footnotesize\ttfamily template$<$typename tuple\+\_\+type $>$ \\
bool \hyperlink{structtuple__less__comparator__1st}{tuple\+\_\+less\+\_\+comparator\+\_\+1st}$<$ tuple\+\_\+type $>$\+::operator() (\begin{DoxyParamCaption}\item[{const tuple\+\_\+type \&}]{\+\_\+a,  }\item[{const tuple\+\_\+type \&}]{\+\_\+b }\end{DoxyParamCaption}) const\hspace{0.3cm}{\ttfamily [inline]}}

The documentation for this struct was generated from the following file\+:\begin{DoxyCompactItemize}
\item 
code/src/common/\hyperlink{widget_8h}{widget.\+h}\end{DoxyCompactItemize}