### 2017-01-24 archive/latex/test_8h.tex

\hypertarget{test_8h}{}\section{code/src/test.h File Reference}
\label{test_8h}\index{code/src/test.\+h@{code/src/test.\+h}}
{\ttfamily \#include \char`\"{}common/common.\+h\char`\"{}}\newline
{\ttfamily \#include \char`\"{}common/tuples.\+h\char`\"{}}\newline
{\ttfamily \#include \char`\"{}common/widget.\+h\char`\"{}}\newline
{\ttfamily \#include \char`\"{}common/basicio.\+h\char`\"{}}\newline
\subsection*{Classes}
\begin{DoxyCompactItemize}
\item 
class \hyperlink{class_test}{Test$<$ alphabet\+\_\+type, alphabet\+\_\+extension\+\_\+type, size\+\_\+type $>$}
\end{DoxyCompactItemize}
\subsection*{Macros}
\begin{DoxyCompactItemize}
\item 
\#define \hyperlink{test_8h_aa8de6ad987cfa13b647db341b8dcc510}{\+\_\+\+T\+E\+S\+T\+\_\+H}
\end{DoxyCompactItemize}

\subsection{Macro Definition Documentation}
\mbox{\Hypertarget{test_8h_aa8de6ad987cfa13b647db341b8dcc510}\label{test_8h_aa8de6ad987cfa13b647db341b8dcc510}} 
\index{test.\+h@{test.\+h}!\+\_\+\+T\+E\+S\+T\+\_\+H@{\+\_\+\+T\+E\+S\+T\+\_\+H}}
\index{\+\_\+\+T\+E\+S\+T\+\_\+H@{\+\_\+\+T\+E\+S\+T\+\_\+H}!test.\+h@{test.\+h}}
\subsubsection{\texorpdfstring{\+\_\+\+T\+E\+S\+T\+\_\+H}{\_TEST\_H}}
{\footnotesize\ttfamily \#define \+\_\+\+T\+E\+S\+T\+\_\+H}