### 2017-01-24 archive/latex/tuples_8h.tex

\hypertarget{tuples_8h}{}\section{code/src/common/tuples.h File Reference}
\label{tuples_8h}\index{code/src/common/tuples.\+h@{code/src/common/tuples.\+h}}
\subsection*{Classes}
\begin{DoxyCompactItemize}
\item 
struct \hyperlink{structpair}{pair$<$ T1, T2 $>$}
\begin{DoxyCompactList}\small\item\em structure pair \end{DoxyCompactList}\item 
struct \hyperlink{structtriple}{triple$<$ T1, T2, T3 $>$}
\begin{DoxyCompactList}\small\item\em structure triple \end{DoxyCompactList}\item 
struct \hyperlink{structquadruple}{quadruple$<$ T1, T2, T3, T4 $>$}
\begin{DoxyCompactList}\small\item\em structure quadruple \end{DoxyCompactList}\end{DoxyCompactItemize}
\subsection*{Functions}
\begin{DoxyCompactItemize}
\item 
{\footnotesize template$<$typename T1 , typename T2 $>$ }\\struct \hyperlink{structpair}{pair} \hyperlink{tuples_8h_a3287af29b6ef2cbe75f1caba87bdde63}{\+\_\+\+\_\+attribute\+\_\+\+\_\+} ((packed))
\item 
\hyperlink{tuples_8h_a861a18bf6751dab7e644b4c4858d3548}{pair} ()
\begin{DoxyCompactList}\small\item\em constructor, default \end{DoxyCompactList}\item 
\hyperlink{tuples_8h_ad95635b05637a43806af080d9ec581ba}{pair} (const \hyperlink{structpair}{pair} \&\+\_\+item)
\begin{DoxyCompactList}\small\item\em constructor, copy \end{DoxyCompactList}\item 
\hyperlink{tuples_8h_a80f1b898df1006722612d5ecf89cd4d7}{pair} (const T1 \&\+\_\+first, const T2 \&\+\_\+second)
\begin{DoxyCompactList}\small\item\em constructor, component \end{DoxyCompactList}\item 
static \hyperlink{structpair}{pair} \& \hyperlink{tuples_8h_ac8a41d1a11f39ac2d9916b2248d76d5d}{min\+\_\+value} ()
\begin{DoxyCompactList}\small\item\em min value \end{DoxyCompactList}\item 
static \hyperlink{structpair}{pair} \& \hyperlink{tuples_8h_abc17aed069c35187f4d6f8b05c506300}{max\+\_\+value} ()
\begin{DoxyCompactList}\small\item\em max vlaue \end{DoxyCompactList}\item 
std\+::ostream \& \hyperlink{tuples_8h_a068f246e1ab74655cd80c5d32b347a20}{operator$<$$<$} (std\+::ostream \+\_\+os, const \hyperlink{structpair}{pair} \&\+\_\+item)
\begin{DoxyCompactList}\small\item\em output \end{DoxyCompactList}\item 
{\footnotesize template$<$typename T1 , typename T2 , typename T3 $>$ }\\struct \hyperlink{structtriple}{triple} \hyperlink{tuples_8h_a61f5f588c8a34e43c27971a61e3a357d}{\+\_\+\+\_\+attribute\+\_\+\+\_\+} ((packed))
\item 
\hyperlink{tuples_8h_a72bccb6f22ab02a7f169a9d9b0671646}{triple} ()
\begin{DoxyCompactList}\small\item\em constructor, default \end{DoxyCompactList}\item 
\hyperlink{tuples_8h_a43fbd7271bb2ea578cbcc6b90491ef12}{triple} (const \hyperlink{structtriple}{triple} \&\+\_\+item)
\begin{DoxyCompactList}\small\item\em constructor, copy \end{DoxyCompactList}\item 
\hyperlink{tuples_8h_a2be7f6af3a54cfc97c55bc395576db62}{triple} (const T1 \&\+\_\+first, const T2 \&\+\_\+second, const T3 \&\+\_\+third)
\begin{DoxyCompactList}\small\item\em constructor, component \end{DoxyCompactList}\item 
std\+::ostream \& \hyperlink{tuples_8h_ae396d2c152e6415ae24b52df6cdfbe23}{operator$<$$<$} (std\+::ostream \+\_\+os, const \hyperlink{structtriple}{triple} \&\+\_\+item)
\begin{DoxyCompactList}\small\item\em output \end{DoxyCompactList}\item 
{\footnotesize template$<$typename T1 , typename T2 , typename T3 , typename T4 $>$ }\\struct \hyperlink{structquadruple}{quadruple} \hyperlink{tuples_8h_aa434f23135e2672ad8fe4ec768f3d88c}{\+\_\+\+\_\+attribute\+\_\+\+\_\+} ((packed))
\item 
\hyperlink{tuples_8h_abc5517aabc587ce4dbe5d5d5c51e926e}{quadruple} ()
\begin{DoxyCompactList}\small\item\em constructor, default \end{DoxyCompactList}\item 
\hyperlink{tuples_8h_ac6ecb9121ff7c7bc8f2bb354d86a7da9}{quadruple} (const \hyperlink{structquadruple}{quadruple} \&\+\_\+item)
\begin{DoxyCompactList}\small\item\em constructor, copy \end{DoxyCompactList}\item 
\hyperlink{tuples_8h_a275ae5ff396331c5f894cae5e717aafd}{quadruple} (const T1 \&\+\_\+first, const T2 \&\+\_\+second, const T3 \&\+\_\+third, const T4 \&\+\_\+forth)
\begin{DoxyCompactList}\small\item\em constructor, component \end{DoxyCompactList}\item 
std\+::ostream \& \hyperlink{tuples_8h_a5ab8584e7c3f4d217350802e27dd64da}{operator$<$$<$} (std\+::ostream \+\_\+os, const \hyperlink{structquadruple}{quadruple} \&\+\_\+item)
\begin{DoxyCompactList}\small\item\em output \end{DoxyCompactList}\end{DoxyCompactItemize}
\subsection*{Variables}
\begin{DoxyCompactItemize}
\item 
T1 \hyperlink{tuples_8h_ae885a6822c40e26649c9dd8b476942f8}{first}
\item 
T2 \hyperlink{tuples_8h_a9a8a97f32d498c09c0f1f82c8a67bd6f}{second}
\item 
T3 \hyperlink{tuples_8h_ab02b336e2ad870f773affd07b0f68ef8}{third}
\item 
T4 \hyperlink{tuples_8h_a9c5c830dde97573ef7be2d18f35d61fd}{forth}
\end{DoxyCompactItemize}

\subsection{Function Documentation}
\mbox{\Hypertarget{tuples_8h_a3287af29b6ef2cbe75f1caba87bdde63}\label{tuples_8h_a3287af29b6ef2cbe75f1caba87bdde63}} 
\index{tuples.\+h@{tuples.\+h}!\+\_\+\+\_\+attribute\+\_\+\+\_\+@{\+\_\+\+\_\+attribute\+\_\+\+\_\+}}
\index{\+\_\+\+\_\+attribute\+\_\+\+\_\+@{\+\_\+\+\_\+attribute\+\_\+\+\_\+}!tuples.\+h@{tuples.\+h}}
\subsubsection{\texorpdfstring{\+\_\+\+\_\+attribute\+\_\+\+\_\+()}{\_\_attribute\_\_()}\hspace{0.1cm}{\footnotesize\ttfamily [1/3]}}
{\footnotesize\ttfamily template$<$typename T1 , typename T2 $>$ \\
struct \hyperlink{structpair}{pair} \+\_\+\+\_\+attribute\+\_\+\+\_\+ (\begin{DoxyParamCaption}\item[{(packed)}]{ }\end{DoxyParamCaption})}

\mbox{\Hypertarget{tuples_8h_a61f5f588c8a34e43c27971a61e3a357d}\label{tuples_8h_a61f5f588c8a34e43c27971a61e3a357d}} 
\index{tuples.\+h@{tuples.\+h}!\+\_\+\+\_\+attribute\+\_\+\+\_\+@{\+\_\+\+\_\+attribute\+\_\+\+\_\+}}
\index{\+\_\+\+\_\+attribute\+\_\+\+\_\+@{\+\_\+\+\_\+attribute\+\_\+\+\_\+}!tuples.\+h@{tuples.\+h}}
\subsubsection{\texorpdfstring{\+\_\+\+\_\+attribute\+\_\+\+\_\+()}{\_\_attribute\_\_()}\hspace{0.1cm}{\footnotesize\ttfamily [2/3]}}
{\footnotesize\ttfamily template$<$typename T1 , typename T2 , typename T3 $>$ \\
struct \hyperlink{structtriple}{triple} \+\_\+\+\_\+attribute\+\_\+\+\_\+ (\begin{DoxyParamCaption}\item[{(packed)}]{ }\end{DoxyParamCaption})}

\mbox{\Hypertarget{tuples_8h_aa434f23135e2672ad8fe4ec768f3d88c}\label{tuples_8h_aa434f23135e2672ad8fe4ec768f3d88c}} 
\index{tuples.\+h@{tuples.\+h}!\+\_\+\+\_\+attribute\+\_\+\+\_\+@{\+\_\+\+\_\+attribute\+\_\+\+\_\+}}
\index{\+\_\+\+\_\+attribute\+\_\+\+\_\+@{\+\_\+\+\_\+attribute\+\_\+\+\_\+}!tuples.\+h@{tuples.\+h}}
\subsubsection{\texorpdfstring{\+\_\+\+\_\+attribute\+\_\+\+\_\+()}{\_\_attribute\_\_()}\hspace{0.1cm}{\footnotesize\ttfamily [3/3]}}
{\footnotesize\ttfamily template$<$typename T1 , typename T2 , typename T3 , typename T4 $>$ \\
struct \hyperlink{structquadruple}{quadruple} \+\_\+\+\_\+attribute\+\_\+\+\_\+ (\begin{DoxyParamCaption}\item[{(packed)}]{ }\end{DoxyParamCaption})}

\mbox{\Hypertarget{tuples_8h_abc17aed069c35187f4d6f8b05c506300}\label{tuples_8h_abc17aed069c35187f4d6f8b05c506300}} 
\index{tuples.\+h@{tuples.\+h}!max\+\_\+value@{max\+\_\+value}}
\index{max\+\_\+value@{max\+\_\+value}!tuples.\+h@{tuples.\+h}}
\subsubsection{\texorpdfstring{max\+\_\+value()}{max\_value()}}
{\footnotesize\ttfamily static \hyperlink{structquadruple}{quadruple} \& max\+\_\+value (\begin{DoxyParamCaption}{ }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [static]}}

max vlaue

max value \mbox{\Hypertarget{tuples_8h_ac8a41d1a11f39ac2d9916b2248d76d5d}\label{tuples_8h_ac8a41d1a11f39ac2d9916b2248d76d5d}} 
\index{tuples.\+h@{tuples.\+h}!min\+\_\+value@{min\+\_\+value}}
\index{min\+\_\+value@{min\+\_\+value}!tuples.\+h@{tuples.\+h}}
\subsubsection{\texorpdfstring{min\+\_\+value()}{min\_value()}}
{\footnotesize\ttfamily static \hyperlink{structquadruple}{quadruple} \& min\+\_\+value (\begin{DoxyParamCaption}{ }\end{DoxyParamCaption})\hspace{0.3cm}{\ttfamily [static]}}

min value

\mbox{\Hypertarget{tuples_8h_a068f246e1ab74655cd80c5d32b347a20}\label{tuples_8h_a068f246e1ab74655cd80c5d32b347a20}} 
\index{tuples.\+h@{tuples.\+h}!operator$<$$<$@{operator$<$$<$}}
\index{operator$<$$<$@{operator$<$$<$}!tuples.\+h@{tuples.\+h}}
\subsubsection{\texorpdfstring{operator$<$$<$()}{operator<<()}\hspace{0.1cm}{\footnotesize\ttfamily [1/3]}}
{\footnotesize\ttfamily friend std\+::ostream\& \+\_\+\+\_\+attribute\+\_\+\+\_\+\+::operator$<$$<$ (\begin{DoxyParamCaption}\item[{std\+::ostream}]{\+\_\+os,  }\item[{const \hyperlink{structpair}{pair} \&}]{\+\_\+item }\end{DoxyParamCaption})}

output

\mbox{\Hypertarget{tuples_8h_ae396d2c152e6415ae24b52df6cdfbe23}\label{tuples_8h_ae396d2c152e6415ae24b52df6cdfbe23}} 
\index{tuples.\+h@{tuples.\+h}!operator$<$$<$@{operator$<$$<$}}
\index{operator$<$$<$@{operator$<$$<$}!tuples.\+h@{tuples.\+h}}
\subsubsection{\texorpdfstring{operator$<$$<$()}{operator<<()}\hspace{0.1cm}{\footnotesize\ttfamily [2/3]}}
{\footnotesize\ttfamily friend std\+::ostream\& \+\_\+\+\_\+attribute\+\_\+\+\_\+\+::operator$<$$<$ (\begin{DoxyParamCaption}\item[{std\+::ostream}]{\+\_\+os,  }\item[{const \hyperlink{structtriple}{triple} \&}]{\+\_\+item }\end{DoxyParamCaption})}

output

\mbox{\Hypertarget{tuples_8h_a5ab8584e7c3f4d217350802e27dd64da}\label{tuples_8h_a5ab8584e7c3f4d217350802e27dd64da}} 
\index{tuples.\+h@{tuples.\+h}!operator$<$$<$@{operator$<$$<$}}
\index{operator$<$$<$@{operator$<$$<$}!tuples.\+h@{tuples.\+h}}
\subsubsection{\texorpdfstring{operator$<$$<$()}{operator<<()}\hspace{0.1cm}{\footnotesize\ttfamily [3/3]}}
{\footnotesize\ttfamily friend std\+::ostream\& \+\_\+\+\_\+attribute\+\_\+\+\_\+\+::operator$<$$<$ (\begin{DoxyParamCaption}\item[{std\+::ostream}]{\+\_\+os,  }\item[{const \hyperlink{structquadruple}{quadruple} \&}]{\+\_\+item }\end{DoxyParamCaption})}

output

\mbox{\Hypertarget{tuples_8h_a861a18bf6751dab7e644b4c4858d3548}\label{tuples_8h_a861a18bf6751dab7e644b4c4858d3548}} 
\index{tuples.\+h@{tuples.\+h}!pair@{pair}}
\index{pair@{pair}!tuples.\+h@{tuples.\+h}}
\subsubsection{\texorpdfstring{pair()}{pair()}\hspace{0.1cm}{\footnotesize\ttfamily [1/3]}}
{\footnotesize\ttfamily \+\_\+\+\_\+attribute\+\_\+\+\_\+\+::pair (\begin{DoxyParamCaption}{ }\end{DoxyParamCaption})}

constructor, default

\mbox{\Hypertarget{tuples_8h_ad95635b05637a43806af080d9ec581ba}\label{tuples_8h_ad95635b05637a43806af080d9ec581ba}} 
\index{tuples.\+h@{tuples.\+h}!pair@{pair}}
\index{pair@{pair}!tuples.\+h@{tuples.\+h}}
\subsubsection{\texorpdfstring{pair()}{pair()}\hspace{0.1cm}{\footnotesize\ttfamily [2/3]}}
{\footnotesize\ttfamily \+\_\+\+\_\+attribute\+\_\+\+\_\+\+::pair (\begin{DoxyParamCaption}\item[{const \hyperlink{structpair}{pair} \&}]{\+\_\+item }\end{DoxyParamCaption})}

constructor, copy

\mbox{\Hypertarget{tuples_8h_a80f1b898df1006722612d5ecf89cd4d7}\label{tuples_8h_a80f1b898df1006722612d5ecf89cd4d7}} 
\index{tuples.\+h@{tuples.\+h}!pair@{pair}}
\index{pair@{pair}!tuples.\+h@{tuples.\+h}}
\subsubsection{\texorpdfstring{pair()}{pair()}\hspace{0.1cm}{\footnotesize\ttfamily [3/3]}}
{\footnotesize\ttfamily \+\_\+\+\_\+attribute\+\_\+\+\_\+\+::pair (\begin{DoxyParamCaption}\item[{const T1 \&}]{\+\_\+first,  }\item[{const T2 \&}]{\+\_\+second }\end{DoxyParamCaption})}

constructor, component

\mbox{\Hypertarget{tuples_8h_abc5517aabc587ce4dbe5d5d5c51e926e}\label{tuples_8h_abc5517aabc587ce4dbe5d5d5c51e926e}} 
\index{tuples.\+h@{tuples.\+h}!quadruple@{quadruple}}
\index{quadruple@{quadruple}!tuples.\+h@{tuples.\+h}}
\subsubsection{\texorpdfstring{quadruple()}{quadruple()}\hspace{0.1cm}{\footnotesize\ttfamily [1/3]}}
{\footnotesize\ttfamily \+\_\+\+\_\+attribute\+\_\+\+\_\+\+::quadruple (\begin{DoxyParamCaption}{ }\end{DoxyParamCaption})}

constructor, default

\mbox{\Hypertarget{tuples_8h_ac6ecb9121ff7c7bc8f2bb354d86a7da9}\label{tuples_8h_ac6ecb9121ff7c7bc8f2bb354d86a7da9}} 
\index{tuples.\+h@{tuples.\+h}!quadruple@{quadruple}}
\index{quadruple@{quadruple}!tuples.\+h@{tuples.\+h}}
\subsubsection{\texorpdfstring{quadruple()}{quadruple()}\hspace{0.1cm}{\footnotesize\ttfamily [2/3]}}
{\footnotesize\ttfamily \+\_\+\+\_\+attribute\+\_\+\+\_\+\+::quadruple (\begin{DoxyParamCaption}\item[{const \hyperlink{structquadruple}{quadruple} \&}]{\+\_\+item }\end{DoxyParamCaption})}

constructor, copy

\mbox{\Hypertarget{tuples_8h_a275ae5ff396331c5f894cae5e717aafd}\label{tuples_8h_a275ae5ff396331c5f894cae5e717aafd}} 
\index{tuples.\+h@{tuples.\+h}!quadruple@{quadruple}}
\index{quadruple@{quadruple}!tuples.\+h@{tuples.\+h}}
\subsubsection{\texorpdfstring{quadruple()}{quadruple()}\hspace{0.1cm}{\footnotesize\ttfamily [3/3]}}
{\footnotesize\ttfamily \+\_\+\+\_\+attribute\+\_\+\+\_\+\+::quadruple (\begin{DoxyParamCaption}\item[{const T1 \&}]{\+\_\+first,  }\item[{const T2 \&}]{\+\_\+second,  }\item[{const T3 \&}]{\+\_\+third,  }\item[{const T4 \&}]{\+\_\+forth }\end{DoxyParamCaption})}

constructor, component

\mbox{\Hypertarget{tuples_8h_a72bccb6f22ab02a7f169a9d9b0671646}\label{tuples_8h_a72bccb6f22ab02a7f169a9d9b0671646}} 
\index{tuples.\+h@{tuples.\+h}!triple@{triple}}
\index{triple@{triple}!tuples.\+h@{tuples.\+h}}
\subsubsection{\texorpdfstring{triple()}{triple()}\hspace{0.1cm}{\footnotesize\ttfamily [1/3]}}
{\footnotesize\ttfamily \+\_\+\+\_\+attribute\+\_\+\+\_\+\+::triple (\begin{DoxyParamCaption}{ }\end{DoxyParamCaption})}

constructor, default

\mbox{\Hypertarget{tuples_8h_a43fbd7271bb2ea578cbcc6b90491ef12}\label{tuples_8h_a43fbd7271bb2ea578cbcc6b90491ef12}} 
\index{tuples.\+h@{tuples.\+h}!triple@{triple}}
\index{triple@{triple}!tuples.\+h@{tuples.\+h}}
\subsubsection{\texorpdfstring{triple()}{triple()}\hspace{0.1cm}{\footnotesize\ttfamily [2/3]}}
{\footnotesize\ttfamily \+\_\+\+\_\+attribute\+\_\+\+\_\+\+::triple (\begin{DoxyParamCaption}\item[{const \hyperlink{structtriple}{triple} \&}]{\+\_\+item }\end{DoxyParamCaption})}

constructor, copy

\mbox{\Hypertarget{tuples_8h_a2be7f6af3a54cfc97c55bc395576db62}\label{tuples_8h_a2be7f6af3a54cfc97c55bc395576db62}} 
\index{tuples.\+h@{tuples.\+h}!triple@{triple}}
\index{triple@{triple}!tuples.\+h@{tuples.\+h}}
\subsubsection{\texorpdfstring{triple()}{triple()}\hspace{0.1cm}{\footnotesize\ttfamily [3/3]}}
{\footnotesize\ttfamily \+\_\+\+\_\+attribute\+\_\+\+\_\+\+::triple (\begin{DoxyParamCaption}\item[{const T1 \&}]{\+\_\+first,  }\item[{const T2 \&}]{\+\_\+second,  }\item[{const T3 \&}]{\+\_\+third }\end{DoxyParamCaption})}

constructor, component

\subsection{Variable Documentation}
\mbox{\Hypertarget{tuples_8h_ae885a6822c40e26649c9dd8b476942f8}\label{tuples_8h_ae885a6822c40e26649c9dd8b476942f8}} 
\index{tuples.\+h@{tuples.\+h}!first@{first}}
\index{first@{first}!tuples.\+h@{tuples.\+h}}
\subsubsection{\texorpdfstring{first}{first}}
{\footnotesize\ttfamily T1 first}

\mbox{\Hypertarget{tuples_8h_a9c5c830dde97573ef7be2d18f35d61fd}\label{tuples_8h_a9c5c830dde97573ef7be2d18f35d61fd}} 
\index{tuples.\+h@{tuples.\+h}!forth@{forth}}
\index{forth@{forth}!tuples.\+h@{tuples.\+h}}
\subsubsection{\texorpdfstring{forth}{forth}}
{\footnotesize\ttfamily T4 forth}

\mbox{\Hypertarget{tuples_8h_a9a8a97f32d498c09c0f1f82c8a67bd6f}\label{tuples_8h_a9a8a97f32d498c09c0f1f82c8a67bd6f}} 
\index{tuples.\+h@{tuples.\+h}!second@{second}}
\index{second@{second}!tuples.\+h@{tuples.\+h}}
\subsubsection{\texorpdfstring{second}{second}}
{\footnotesize\ttfamily T2 second}

\mbox{\Hypertarget{tuples_8h_ab02b336e2ad870f773affd07b0f68ef8}\label{tuples_8h_ab02b336e2ad870f773affd07b0f68ef8}} 
\index{tuples.\+h@{tuples.\+h}!third@{third}}
\index{third@{third}!tuples.\+h@{tuples.\+h}}
\subsubsection{\texorpdfstring{third}{third}}
{\footnotesize\ttfamily T3 third}