### 2017-01-24 archive/latex/validate2_8cpp.tex

\hypertarget{validate2_8cpp}{}\section{code/src/validate2.cpp File Reference}
\label{validate2_8cpp}\index{code/src/validate2.\+cpp@{code/src/validate2.\+cpp}}
{\ttfamily \#include \char`\"{}validate2.\+h\char`\"{}}\newline
\subsection*{Functions}
\begin{DoxyCompactItemize}
\item 
int \hyperlink{validate2_8cpp_a3c04138a5bfe5d72780bb7e82a18e627}{main} (int argc, char $\ast$$\ast$argv)
\end{DoxyCompactItemize}

\subsection{Function Documentation}
\mbox{\Hypertarget{validate2_8cpp_a3c04138a5bfe5d72780bb7e82a18e627}\label{validate2_8cpp_a3c04138a5bfe5d72780bb7e82a18e627}} 
\index{validate2.\+cpp@{validate2.\+cpp}!main@{main}}
\index{main@{main}!validate2.\+cpp@{validate2.\+cpp}}
\subsubsection{\texorpdfstring{main()}{main()}}
{\footnotesize\ttfamily int main (\begin{DoxyParamCaption}\item[{int}]{argc,  }\item[{char $\ast$$\ast$}]{argv }\end{DoxyParamCaption})}