### 2017-01-24 archive/latex/validate2_8h.tex

\hypertarget{validate2_8h}{}\section{code/src/validate2.h File Reference}
\label{validate2_8h}\index{code/src/validate2.\+h@{code/src/validate2.\+h}}

validate sa and lcp using internal memory sorts and Karp-\/rabin finger-\/print function.

{\ttfamily \#include \char`\"{}common/common.\+h\char`\"{}}\newline
{\ttfamily \#include \char`\"{}common/tuples.\+h\char`\"{}}\newline
{\ttfamily \#include \char`\"{}common/widget.\+h\char`\"{}}\newline
{\ttfamily \#include \char`\"{}common/basicio.\+h\char`\"{}}\newline
\subsection*{Classes}
\begin{DoxyCompactItemize}
\item 
class \hyperlink{class_validate2}{Validate2$<$ alphabet\+\_\+type, size\+\_\+type $>$}
\begin{DoxyCompactList}\small\item\em a suffix and L\+CP array validater \end{DoxyCompactList}\item 
struct \hyperlink{struct_validate2_1_1_r_interval}{Validate2$<$ alphabet\+\_\+type, size\+\_\+type $>$\+::\+R\+Interval}
\begin{DoxyCompactList}\small\item\em store (R \% P)$^\wedge$1, (R \% P)$^\wedge$2, (R \% P)$^\wedge$4, ... \end{DoxyCompactList}\end{DoxyCompactItemize}

\subsection{Detailed Description}
validate sa and lcp using internal memory sorts and Karp-\/rabin finger-\/print function.

Copyright (c) 2016, Sun Yat-\/sen University, All rights reserved

\begin{DoxyNote}{Note}
The alphaber size of the input string is assumed to be O(1) or O(n).

This method requires multiple sequential scans for the input string. During each scan, all the fingerprints are computed by an iterative way. This has become the bottleneck to the system\textquotesingle{}s running time.
\end{DoxyNote}
\begin{DoxyAuthor}{Author}
Yi Wu 
\end{DoxyAuthor}
\begin{DoxyDate}{Date}
2016.\+12 
\end{DoxyDate}