### 2017-01-24 archive/latex/validate3_8cpp.tex

\hypertarget{validate3_8cpp}{}\section{code/src/validate3.cpp File Reference}
\label{validate3_8cpp}\index{code/src/validate3.\+cpp@{code/src/validate3.\+cpp}}
{\ttfamily \#include \char`\"{}validate3.\+h\char`\"{}}\newline
\subsection*{Functions}
\begin{DoxyCompactItemize}
\item 
int \hyperlink{validate3_8cpp_a3c04138a5bfe5d72780bb7e82a18e627}{main} (int argc, char $\ast$$\ast$argv)
\end{DoxyCompactItemize}
\subsection*{Variables}
\begin{DoxyCompactItemize}
\item 
char $\ast$ \hyperlink{validate3_8cpp_a598d558168ea9fcc902baf4a3a42a5ea}{prog\+\_\+name}
\end{DoxyCompactItemize}

\subsection{Function Documentation}
\mbox{\Hypertarget{validate3_8cpp_a3c04138a5bfe5d72780bb7e82a18e627}\label{validate3_8cpp_a3c04138a5bfe5d72780bb7e82a18e627}} 
\index{validate3.\+cpp@{validate3.\+cpp}!main@{main}}
\index{main@{main}!validate3.\+cpp@{validate3.\+cpp}}
\subsubsection{\texorpdfstring{main()}{main()}}
{\footnotesize\ttfamily int main (\begin{DoxyParamCaption}\item[{int}]{argc,  }\item[{char $\ast$$\ast$}]{argv }\end{DoxyParamCaption})}

\subsection{Variable Documentation}
\mbox{\Hypertarget{validate3_8cpp_a598d558168ea9fcc902baf4a3a42a5ea}\label{validate3_8cpp_a598d558168ea9fcc902baf4a3a42a5ea}} 
\index{validate3.\+cpp@{validate3.\+cpp}!prog\+\_\+name@{prog\+\_\+name}}
\index{prog\+\_\+name@{prog\+\_\+name}!validate3.\+cpp@{validate3.\+cpp}}
\subsubsection{\texorpdfstring{prog\+\_\+name}{prog\_name}}
{\footnotesize\ttfamily char$\ast$ prog\+\_\+name}