### 2017-01-24 archive/latex/validate3_8h.tex

\hypertarget{validate3_8h}{}\section{code/src/validate3.h File Reference}
\label{validate3_8h}\index{code/src/validate3.\+h@{code/src/validate3.\+h}}

validate sa and lcp using external memory sorts and Karp-\/rabin finger-\/print function.

{\ttfamily \#include \char`\"{}common/common.\+h\char`\"{}}\newline
{\ttfamily \#include \char`\"{}common/tuples.\+h\char`\"{}}\newline
{\ttfamily \#include \char`\"{}common/widget.\+h\char`\"{}}\newline
{\ttfamily \#include \char`\"{}common/basicio.\+h\char`\"{}}\newline
\subsection*{Classes}
\begin{DoxyCompactItemize}
\item 
class \hyperlink{class_validate3}{Validate3$<$ alphabet\+\_\+type, size\+\_\+type $>$}
\begin{DoxyCompactList}\small\item\em validate sa and lcp using Karp-\/\+Rabin fingerprinting function \end{DoxyCompactList}\item 
struct \hyperlink{struct_validate3_1_1_r_interval}{Validate3$<$ alphabet\+\_\+type, size\+\_\+type $>$\+::\+R\+Interval}
\begin{DoxyCompactList}\small\item\em store (R \% P)$^\wedge$1, (R \% P)$^\wedge$2, (R \% P)$^\wedge$4, ... \end{DoxyCompactList}\end{DoxyCompactItemize}
\subsection*{Macros}
\begin{DoxyCompactItemize}
\item 
\#define \hyperlink{validate3_8h_aed71fcd0df9f567ccbffeda93549cefc}{T\+E\+S\+T\+\_\+\+V\+A\+L\+I\+D\+A\+T\+E3}
\end{DoxyCompactItemize}

\subsection{Detailed Description}
validate sa and lcp using external memory sorts and Karp-\/rabin finger-\/print function.

Copyright (c) 2016, Sun Yat-\/sen University, All rights reserved

\begin{DoxyNote}{Note}
The alphaber size of the input string is assumed to be O(1) or O(n).

This is the first method reported in xxx.
\end{DoxyNote}
\begin{DoxyAuthor}{Author}
Yi Wu 
\end{DoxyAuthor}
\begin{DoxyDate}{Date}
2016.\+12 
\end{DoxyDate}

\subsection{Macro Definition Documentation}
\mbox{\Hypertarget{validate3_8h_aed71fcd0df9f567ccbffeda93549cefc}\label{validate3_8h_aed71fcd0df9f567ccbffeda93549cefc}} 
\index{validate3.\+h@{validate3.\+h}!T\+E\+S\+T\+\_\+\+V\+A\+L\+I\+D\+A\+T\+E3@{T\+E\+S\+T\+\_\+\+V\+A\+L\+I\+D\+A\+T\+E3}}
\index{T\+E\+S\+T\+\_\+\+V\+A\+L\+I\+D\+A\+T\+E3@{T\+E\+S\+T\+\_\+\+V\+A\+L\+I\+D\+A\+T\+E3}!validate3.\+h@{validate3.\+h}}
\subsubsection{\texorpdfstring{T\+E\+S\+T\+\_\+\+V\+A\+L\+I\+D\+A\+T\+E3}{TEST\_VALIDATE3}}
{\footnotesize\ttfamily \#define T\+E\+S\+T\+\_\+\+V\+A\+L\+I\+D\+A\+T\+E3}