### 2017-01-24 archive/latex/validate_8h.tex

\hypertarget{validate_8h}{}\section{code/src/validate.h File Reference}
\label{validate_8h}\index{code/src/validate.\+h@{code/src/validate.\+h}}

first method for validating suffix and L\+CP arrays

{\ttfamily \#include \char`\"{}common/common.\+h\char`\"{}}\newline
{\ttfamily \#include \char`\"{}common/basicio.\+h\char`\"{}}\newline
{\ttfamily \#include \char`\"{}common/widget.\+h\char`\"{}}\newline
{\ttfamily \#include \char`\"{}stxxl/timer\char`\"{}}\newline
\subsection*{Classes}
\begin{DoxyCompactItemize}
\item 
class \hyperlink{class_validate}{Validate$<$ alphabet\+\_\+type, size\+\_\+type, offset\+\_\+type $>$}
\begin{DoxyCompactList}\small\item\em validate suffix and L\+CP arrays using Karp-\/\+Rarbin finger-\/printing function \end{DoxyCompactList}\item 
struct \hyperlink{struct_validate_1_1_r_interval}{Validate$<$ alphabet\+\_\+type, size\+\_\+type, offset\+\_\+type $>$\+::\+R\+Interval}
\begin{DoxyCompactList}\small\item\em store (R \% P)$^\wedge$1, (R \% P)$^\wedge$2, (R \% P)$^\wedge$4, ... \end{DoxyCompactList}\end{DoxyCompactItemize}
\subsection*{Macros}
\begin{DoxyCompactItemize}
\item 
\#define \hyperlink{validate_8h_a864a52c772f23a04bcef46a40c9efed4}{T\+E\+S\+T\+\_\+1}
\item 
\#define \hyperlink{validate_8h_a1cc15202a800a32df74352e452309c57}{S\+T\+A\+T\+I\+S\+T\+I\+C\+S\+\_\+\+C\+O\+L\+L\+E\+C\+T\+I\+ON}
\end{DoxyCompactItemize}

\subsection{Detailed Description}
first method for validating suffix and L\+CP arrays

Copyright (c) 2016, Sun Yat-\/sen University, All rights reserved

Given suffix and L\+CP arrays, validate their correctness using external memory sorts and Karp-\/\+Rabin Finger-\/printing function.

The alphabet of input string is assumed to be O(1) or O(n).

\begin{DoxyAuthor}{Author}
Yi Wu 
\end{DoxyAuthor}
\begin{DoxyDate}{Date}
2016.\+12 
\end{DoxyDate}

\subsection{Macro Definition Documentation}
\mbox{\Hypertarget{validate_8h_a1cc15202a800a32df74352e452309c57}\label{validate_8h_a1cc15202a800a32df74352e452309c57}} 
\index{validate.\+h@{validate.\+h}!S\+T\+A\+T\+I\+S\+T\+I\+C\+S\+\_\+\+C\+O\+L\+L\+E\+C\+T\+I\+ON@{S\+T\+A\+T\+I\+S\+T\+I\+C\+S\+\_\+\+C\+O\+L\+L\+E\+C\+T\+I\+ON}}
\index{S\+T\+A\+T\+I\+S\+T\+I\+C\+S\+\_\+\+C\+O\+L\+L\+E\+C\+T\+I\+ON@{S\+T\+A\+T\+I\+S\+T\+I\+C\+S\+\_\+\+C\+O\+L\+L\+E\+C\+T\+I\+ON}!validate.\+h@{validate.\+h}}
\subsubsection{\texorpdfstring{S\+T\+A\+T\+I\+S\+T\+I\+C\+S\+\_\+\+C\+O\+L\+L\+E\+C\+T\+I\+ON}{STATISTICS\_COLLECTION}}
{\footnotesize\ttfamily \#define S\+T\+A\+T\+I\+S\+T\+I\+C\+S\+\_\+\+C\+O\+L\+L\+E\+C\+T\+I\+ON}

\mbox{\Hypertarget{validate_8h_a864a52c772f23a04bcef46a40c9efed4}\label{validate_8h_a864a52c772f23a04bcef46a40c9efed4}} 
\index{validate.\+h@{validate.\+h}!T\+E\+S\+T\+\_\+1@{T\+E\+S\+T\+\_\+1}}
\index{T\+E\+S\+T\+\_\+1@{T\+E\+S\+T\+\_\+1}!validate.\+h@{validate.\+h}}
\subsubsection{\texorpdfstring{T\+E\+S\+T\+\_\+1}{TEST\_1}}
{\footnotesize\ttfamily \#define T\+E\+S\+T\+\_\+1}