### 2017-01-24 archive/latex/widget_8h.tex

\hypertarget{widget_8h}{}\section{code/src/common/widget.h File Reference}
\label{widget_8h}\index{code/src/common/widget.\+h@{code/src/common/widget.\+h}}

sort elements in external memory

{\ttfamily \#include \char`\"{}common.\+h\char`\"{}}\newline
{\ttfamily \#include \char`\"{}stxxl/sorter\char`\"{}}\newline
\subsection*{Classes}
\begin{DoxyCompactItemize}
\item 
struct \hyperlink{struct_pair_less1st}{Pair\+Less1st$<$ pair\+\_\+type $>$}
\begin{DoxyCompactList}\small\item\em sort pair by 1st component in ascending order \end{DoxyCompactList}\item 
struct \hyperlink{struct_pair_less2nd}{Pair\+Less2nd$<$ pair\+\_\+type $>$}
\begin{DoxyCompactList}\small\item\em sort pair by 2nd component in ascending order \end{DoxyCompactList}\item 
struct \hyperlink{struct_ex_vector}{Ex\+Vector$<$ value\+\_\+type $>$}
\begin{DoxyCompactList}\small\item\em template for vector \end{DoxyCompactList}\item 
struct \hyperlink{structtuple__less__comparator__1st}{tuple\+\_\+less\+\_\+comparator\+\_\+1st$<$ tuple\+\_\+type $>$}
\begin{DoxyCompactList}\small\item\em function object for comparing tuples by their first component in ascending order \end{DoxyCompactList}\item 
struct \hyperlink{structtuple__great__comparator__1st}{tuple\+\_\+great\+\_\+comparator\+\_\+1st$<$ tuple\+\_\+type $>$}
\begin{DoxyCompactList}\small\item\em function object for comparing tuples by their first component in ascending order \end{DoxyCompactList}\item 
struct \hyperlink{structtuple__less__comparator__2nd}{tuple\+\_\+less\+\_\+comparator\+\_\+2nd$<$ tuple\+\_\+type $>$}
\begin{DoxyCompactList}\small\item\em function object for comparing tuples by their first two component in ascending order \end{DoxyCompactList}\item 
struct \hyperlink{struct_ex_tuple_sorter}{Ex\+Tuple\+Sorter$<$ tuple\+\_\+type, tuple\+\_\+comparator\+\_\+type $>$}
\begin{DoxyCompactList}\small\item\em template for sorter \end{DoxyCompactList}\end{DoxyCompactItemize}

\subsection{Detailed Description}
sort elements in external memory

Copyright (c) 2016, Sun Yat-\/sen University, All rights reserved

Input elements are grouped into multiple blocks. Elements in each block are sorted in R\+AM and the result is stored in external memory. The block-\/wise results are combined into a whole using a heap in R\+AM at last.

\begin{DoxyAuthor}{Author}
Yi Wu 
\end{DoxyAuthor}
\begin{DoxyDate}{Date}
2016.\+12 
\end{DoxyDate}